### 2009-11-11 docs/marcello-dissertation/thesis/chapter2.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{State of the Art}

Different aspects related to two major fields of Computer Science have
motivated to the inception of this work. In the perspective of Database
Systems and Distributed Systems, this work seeks a different approach for a
persisting collected data from sensor networks into a storage system that can
be seen as a cluster of distributed databases. Likewise, in the perspective of
Software Engineering, this work adds a major capability to an existing sensor 
network framework, meeting the requirements and interfaces. In general, this 
chapter gives the foundation based on the motivations for the inception of a 
different approach to persisting data using a different data model, covering 
topics that will be referred throughout the following other chapters.

This chapter will first give an overview on different related problems
described  in the literature that solved problems related to data modeling and 
representation in sensor networks. Then, it presents the design, architecture 
and implementation details of the DSP Platform, which is the bases for this 
dissertation. Finally, this chapter will briefly give a tutorial on advanced 
topics such as the differences between the KVP data model and the relational 
data model, as well as discussing about the approaches used on Database
Sharding.

\section{Sensor Networks}

Sensor Networks are specialized types of network systems comprised of network 
nodes with specific goals to observe and collect data. The following
subsections describe classes of characteristis of these types of sensor
networks, with applications on research, industrial or militery.

\subsection{Deployment and Mobility}

Each of the nodes can be deployed in a random way such as being dropped
from an airplane, or in a manual fashion at deliberately chosen geographical 
points \cite{snintro01}. Moreover, the deployment can be done as a one-time
event or an iterative way.

However, nodes can be able to move around the deployed area or be static in
this place. In the case of the degree of the former behaviour, a sensor node
can move occasionally or continuously, being a result of an incidental 
side-effect or a desired property of the device, where they are an automotive
device or attached to another moving object. For this reason, \cite{snintro01} 
describes those types of mobility as active or passive.

\subsection{Device Size, Resources, Cost, Energy, Heterogeneity}

Such sensor nodes can be of different sizes, ranging from microscopically small
particles, small devices to bigger ones with a few feet cubic. In addition to
being limited in size, they are also limited in resources and computational
power, forcing their conception limited to have its energy stored in either
batteries or scavenged from solar cells \cite{snintro01}.

Since sensor devices range on size and capabilities, the cost for each of the
devices very from hundreds to thousands of dollars, depending on its
functionality, energy utilization, etc. For thie reason, \cite{snintro01}
explains that a sensor network may potentially contain similar sensor devices,
as they can contain dissimilar ones from each other, which characterizes the heterogeneity 
of a sensor network by the hardware capabilities such as GPS enabled or having
GSM  network connectivity.

\subsection{Communication Modality, Network Infrastructure, Coverage and
Connectivity}
\label{sec:sn-infrastructure}

Sensor nodes can communicate in different modalities such as radio, diffuse
light, inductive or sound, which may influence on the type of protocols used to
communicate among the nodes, where they can be said to be in an organized
infrastructure or as an ad hoc way \cite{snintro01}.

When it comes to the sensor network different organizational infrastructure,
they can differ in types of topology: single-hop, star, networked stars, tree
and graph, where each of them may affect the characteristics of the network as it 
may also influence on data gathering and routing mechanisms. While in a
single-hop sensor network there are only two levels of sensor connectivity, the
tree, graph or networked stars networks uses communications among the network
hosts for data communication.

Last, but not least, sensor networks may be influenced its coverage designed in
its purpose, its size in terms of number of nodes and its relating lifetime. As
sensors are deployed in different areas, they can be distributed in a sparse or
dense way, yet providing necessary coverage for the environment. However, 
considering that sensor devices are very likely to constant failures, projects 
have provided redundant nodes \cite{snintro01}. This degree of coverage also 
influences on the connectivity of the sensor nodes, where they can be
considered connected when a connection is available at all times. On the other
hand, nodes can communicate with other nodes in an intermittent way or sporadic
way. In the former, the sensor node is partitioned, while in the latter the
nodes are isolated most of the time and communicates at certain moments in
time.

\subsection{Network Size, Lifetime and Quality-Of-Service}

Similar to the degree of coverage of the sensor network is the number of
sensor nodes deployed in the network, where they can be in the order of
dozens, hundreds or even thousand of nodes \cite{snintro01}.

The application of the sensor network determines how long the sensor
network may be exist. For example, the lifetime of an environmental sensor
networks \cite{snex01} may be not determined since its service serves in a
regular basis, whereas the lifetime of a sensor network to determine the number
of people in an event is limited to the event.

Finally, the design of sensor networks can be related to the constraints
related to quality-of-service (QoS) such as real-time delivery within certain
period of time or robustness such as maintaining data alive even though there
a communication link is not available \cite{snintro02}.

\section{Data Persistence in Sensor Networks}

The most important purpose of a sensor network is related to its main product:
the use of the raw data. Data collection in sensor networks is directly
dependent on the specifications of the sensor network, its routing mechanisms
and the final destination of the produced data. In a nutshell,
\cite{sn-storage03} describes data as follows:

\begin{itemize}
  \item \textbf{low-level property} as the raw data collected from the
  sensor device; For example, the sea level as a number;
  \item \textbf{Observation or Event} as the processed data in format of the final
  observation. For instance, if the ocean is in its high tight or low tight.
\end{itemize}

\subsection{The Purpose of the Collected Data}
\label{sec:sn-data-purpose}

At first glance, the raw data serves as the primary product of the sensor
network. However, the first consideration that must take place is regarding the
purpose of the collected data, as \cite{sn-provenance01} explains
differentiates between two main use of the sensor data:

\begin{itemize}
  \item \textbf{Real-time Data Stream}, as the generated data from the sensor is
  actually not transmitted, but extracted from the device as it is requested.
  The data is still in its natural format specified by the sensor manufacturer;
  \item \textbf{Archival Data}, as the collected data have historical value or will
  be use by other applications at a future time. 
\end{itemize}

While real-time data stream are used and can potentially be discarded, the
archival data needs to be stored in a secondary memory. Nonetheless, given the
nature of the sensor device specification, the low-level data produced
by a given sensor device might not contain necessary information regarding the
identity of the data itself. The use of Data Provenance techniques is described
by \cite{sn-provenance01} as an important step to provide the collected data 
specific metadata regarding its location and lineage.

Under these circumstances, the collected data needs first to face its jorney
starting from its transmission from its producing device to the specified data 
sink as specified by \cite{sn-storage01} \cite{sn-storage02} \cite{sn-storage03}.

\subsubsection{Providing Identity to the Collected Sensor Data}

Most of sensor networks implementations use the collected data from the sensor
devices after they observe the environment. \cite{sn-provenance01} showed that
collected data used for archival purposes are usually missing important
features. First, the device's sampling are usually in raw format or without
obvious names. In order to provide search over the collected sensor data, the
data needs first to be described and indexed, and this process can be related
to Data Provenance \cite{db-provenance}.

An example of Data Provenance application is the use of aggregation of data
over time to estimate the changing effects of the weather in a given region.
For this reason, the collected data needs descriptions or annotations that can
give enough information to track changes, that is, the history of how and when
a given data has changed. The following should be the foundation of a
provenance-aware collected data according to \cite{sn-provenance01}:

\begin{itemize}
  \item each low-level observation as it describes the actual data;
  \item the location where the collected data was generated;
  \item the time when the data was collected.
\end{itemize}

\subsection{Storage Location for Collected Data}
\label{sec:sn-storage-locations}

Given that collected data can be used for different purposes as described in
section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
responsible to send the raw data to a specified data storage destination. Upon
producing data, a sensor network node is specified to deal with the produced
data, where in most cases, \cite{sn-storage03} described that depending on the
sensor network infrastructure (see section \ref{sec:sn-infrastructure}),
different storage mechanisms are provided:

\begin{itemize}
  \item \textbf{External Storage}: the sensor device sends the produced data to
  another sensor node that contains an external storage. 
  \item \textbf{Local Storage}: as it is the case of real-time data stream, where data
  is allocated on the sensor device's main read-only memory.
\end{itemize}

It is clear that the data flows from the sensor node to another depending on
the specification of the network. On the whole, the sensor network
infrastructure dictates where the data will flow to, as described in section
\ref{sec:sn-infrastructure}. \cite{sn-storage02} sees a wireless sensor
networks as a many-to-one data gathering pattern, when its organization is
based on single-hop or star. On the other hand, when the data organization is
structured as a graph, tree, etc, the data flows from one intermediate sensor
node to another until it reaches its final storage node as described by
\cite{sn-storage01} \cite{sn-storage03}.

Last, but not least, yet a different storage strategy can be used: Data-Centric
Storage. This strategy aims at clustering the collected data based on a given nature of the
data such as location. Regardless the data storage mechanism used, the query
processing will depend on the sensor network infrastructure and data location.

\subsection{Query Processing for Storage Locations}

Given the different locations where the collected data might be as described in
section \ref{sec:sn-storage-locations}, different query processing for the
collected sensor data can be used as described by \cite{sn-storage03}:

\begin{itemize}
  \item {In-Network Query Processing}: sensor nodes that provides real-time
  data stream usually replies to requests to properties. Likewise, when the
  data is used for archival purpose, the data can be located at a network node
  specialized in storing the neighbors collected data. In this way, when a
  query is issued to the sensor network, the query processor will
  collectively verify every storage node for the collected data.
  \item {Centralized Query Processing}: a main network sink receives the
  collected data from the peer nodes. In this case, the query processor queries
  the solo centralized node for the overall values of the participating sensor
  devices.
\end{itemize}

There are pros and cons related to both of the approaches of query processing,
which are also related to the infrastructure, sensor network nature and 
requirements. The main goal of the in-network query processing is to decrease
the amount of data transmitted from the sensor devices to the final data sink, as
sensor devices are usually limited by constraints such as energy utilization 
\cite{sn-storage03}. Along with this strategy, the literature describes
different proposals that aims at upgrading the Structured Query Language (SQL)
adding location-sensitive clauses to the language \cite{sn-db-newop01}
\cite{sn-db-newop02}. However, this main problem related to this approach is
that it may potentially flood all nodes with query messages \cite{sn-storage04}.

In order to mitigate the problems related flooded nodes in in-network query
processing, the use of a Data-Centric storage strategy has been shown to be a
proven alternative for the sake of energy consumption \cite{sn-storage03}
\cite{sn-storage01}. Its main goal is to classify the collected data based in a
specific property, say location or event, in order to be sent to specific
storage nodes that relates to the observation. As the main outcome of this
strategy is to create network partitioning strategy, the query processor would
hit specific sensor nodes with the needed data, and therefore, would
decrease the load and maximize the capacity on the actual sensor devices.

A centralized network query processing processes data in a network node
specified as the network sink. The network sink usually is a network
node with a database server with the collection of data delivered by
all the sensor devices from the network. As a consequence of this
many-to-one communication strategy \cite{sn-storage02}, the query process does
not affect any of the sensor nodes in the network and, thus, maximizes the
sensor device utilization. Moreover, the use of the data for historical
purposes is easier to manage and give access to the primary end users of the
data, following the process of data processing, compression, aggregation, etc
\cite{sn-db02}.

Although this approach seems to be convenient to the point of view of the
sensor nodes and used as the tradeoff between achieving capacity and energy
consumption, the centralized data collector suffers from the annavoiable
creation of a point of traffic concentration \cite{sn-storage02}, also referred
to as the phenomenon of congestion, or funneling effect \cite{sn-storage04}.
For this reason, simple distributed system techniques such as the use of
master-slave or data replication can be used to decrease the overhead.

\section{SF-BEAMS: a Marine Sensor Network for Water Quality Monitoring}

The San Francisco Bay Environmental Assessment and Monitoring Station,
or SF-BEAMS, is a marine sensor network operated by the Romberg Tiburon Center
(RTC). This section details the SF-BEAMS sensor network located at
the Tiburon Island, California, and then presents the details of NetBEAMS, as
it is part of the case study of this work.

\subsection{The SF-BEAMS Infrastructure}

The RTC's main focuses are in the study of complex marine and estuarine
environments using the SF-BEAMS sensors deployed offshore of its pier located
on the banks of the San Francisco Bay at the Tiburon Island, California. In
this way, the project is responsible for providing data for water quality
monitoring, as well as weather and surface conditions.

In general, the SF-BEAMS network infrastructure contains heterogeneous wired and
wireless devices standing on pilons docked in the pier, where each of them is
responsible for observing different conditions of the area, as well as having
its own mechanisms for internal storage for the collected data. In this way,
data can be directly transferred to the labs via ethernet cables or collected
manually download by a staff's laptop computer. Figure \ref{fig:sf-beams}
shows a picture taken from the SF-BEAMS web-cam placed in-site.

\begin{figure}
  \centering
    \includegraphics{../diagrams/cam_image-oct15}
  \caption{Picture of the SF-BEAMS Network, at the RTC pier. Tiburon, CA.}
  \label{fig:sf-beams}
\end{figure}

\subsection{SF-BEAMS produced data and collection process}
\label{sec:sfbeams}

As described in section \ref{sec:sn-infrastructure}, sensor devices generates
its observed data based on properties of measurements defined by its
manufacturer. SF-BEAMS is a single-hop star sensor network containing
different instances of sensors, including the "YSI 6600 ESD V2", as seen on
figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
device that produces around 52 bytes (13x4 Bytes) on a single real-time data
stream as shown on table \ref{tab:ysi-data-stream}.

\begin{figure}
  \centering
  \includegraphics{../diagrams/ysi-device}
  \caption{Picture of the case study sonde device: YSI 6600 ESD V2}
  \label{fig:ysi-device}
\end{figure}

\begin{table}
    \caption{YSI Data Stream}
    \begin{center}
        \begin{tabular}{lr}
  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83 
  0.4     8.7
        \end{tabular}
    \end{center}
    \label{tab:ysi-data-stream}
\end{table}

In order to have access to the observed data, the network staff downloads
the data using one of the device's connection such as the RS-232 serial
connector \cite{rs232}. Then, the data is transferred to the RTC labs for
archival, where it is described, indexed and distributed to its main users.
According to its staff members, in May 2009 the infrastructure of the SF-BEAMS
included the following was organized as follows:

\begin{itemize}
  \item 5 YSI sondes in operation at the RTC site Pier in Tiburon Island, San
  Francisco Bay;
  \item Each YSI sonde produces 52 Bytes for each its observations;
  \item The sampling frequency rate could be configured in ranges of 1, 6 or
  15 minutes, depending on the staff's speficiation. 
\end{itemize}

Considering the infrastructure previously described, the size of the data
produced by the YSI sonde sensors at the RTC pier can estimated by its
measurement frequency as follows:

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Amount of data produced by the RTC's YSI sondes}
        \begin{center}
        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
        \textbf{Weeky} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
        \end{tabular}
        \end{center}
\end{table}

As it is shown, the data produced by this type of sensor runs in the order of
at most hundreds of Megabytes a year, what can be considered a very low data
storage requirement. Upon collecting the data from sensors, the RTC lab staff
use automation scripts written in Matlab to process, index and distribute the
raw data in different formats, including the OPeNDAP \cite{opendap}, a format
widely used in the research institutions that promotes easier data exchange
among them.

Although the execution of the RTC SF-BEAM sensor network produces its
execution, many operational challenges were faced by this type of
infrastructure, when it comes to the data utilization of wireless sensors that
are deployed in offshore conditions, \cite{netbeams2009} described the
operational challenges faced by the RTC staff and proposed NetBEAMS, a
component-based infrastructure that uses common-off-the-shelfs (COTS) embedded
systems and software to automate the process of data collection. Next section
describes how NetBEAMS is used to extract data from the SF-BEAMS sensor network.

\section{NetBEAMS: a Real-World Case Study on Environmental Sensor Network}

The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
is a work-in-progress joint project proposed by the department of Computer
Science and the RTC at San Francisco State University. As described by
\cite{netbeams2009}, the solution can be used to automated the data gathering
process of the SF-BEAMS as described in section \ref{sec:sfbeams}. This
section fully details how NetBEAMS is executed in order to collect data from
SF-BEAMS, describing pertinent design and architectural details.

One of the motivations for the inception of NetBEAMS was the automation
of the data gathering and processing from the SF-BEAMS sondes without requiring
human intervention. Most importantly, the decrease with the costs of constant 
offshore site visits by boat, where biologists need manually collect the data
from the sensor devices in order to take them for further processing at the
lab. For this reason, the NetBEAMS research group proposed the use of the Data
Sensor Platform (DSP) \cite{netbeams2009}, as the second attempt to automate
the operation of SF-BEAMS.

This following sections details the DSP, its architecture and how data is
transported from the network nodes to the main network sink, the RTC labs.

\subsection{The Data Sensor Platform (DSP)}

The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
architecture developed on top of the Java modular framework called OSGi
\cite{osgi}. The platform represents the foundation of the execution of
NetBEAMS, since it is executed from an embedded hardware called Gumstix
\cite{gumstix} using a Linux operating system and on top of the Java Virtual
Machine (JVM). In order to transport data from sensors hosts to the network
sink, the use of a cellular connection is used to transport the collected data
from the senso devices using the HTTP Protocol, as described by
\cite{netbeams09}. Figure \ref{fig:sf-netbeams-node} shows the
architecture of the inhanced node designed by the NetBEAMS research group.

\begin{figure}
  \centering
  \includegraphics{../diagrams/DSP-Gateway-Node}
  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
  \label{fig:sf-netbeams-node}
\end{figure}

The following list describes each of the devices in the NetBEAMS Node.

\begin{itemize}
  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
  responsible for observing the environment. It is connected to the Gumstix
  embedded system, where the DSP is executed;
  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
  provides a computer-on-module environment for the development of small
  embedded systems. This is the main environment of data extraction and remote
  processing is accomplished. It runs a cross-compiled version of Linux Kernel
  2.6;
  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
  Gumstix to the RTC Labs Data Center.
\end{itemize}

The DSP runs inside of the embedded system Gumstix. The following sections
details its design and process for extracting data.

\subsection{The DSP System Design}

This section focus on the development of a Software Platform for the NetBEAMS Gateway Embedded System, or on the system developed in the Gumstix device. As shown on Image 2.4.1.1, the NetBEAMS Gateway Node is an embedded system that contains a set of layer found in any computer. However, each of the resources described from the system were customized for the NetBEAMS system. The main components of such system can be summarized as follows:

    * Operating System: it uses a cross-compiled Gentoo Linux, which supports a wide variety of development tools, including the Java Virtual Machine (JVM), the underlying platform system;
    * Java Virtual Machine: it uses the JamVM, a ~200 KB version of the Sun Microsystems Java Virtual Machine, version 2.0;
    * OSGi Framework: it uses the Knopflerfish implementation of the OSGi 4.1 specification. More details in the following sections.
    * DSP Platform and other Bundles: The plug-and-play DSP Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.