### 2009-11-11 docs/marcello-dissertation/thesis/main.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\chapter*{Appendix A: Triangulations of Polytopes}
%\addcontentsline{toc}{chapter}{Appendix A: Triangulations of Polytopes}

\appendix{Appendix A: Triangulations of Polytopes}

This creates an appendix, which is not numbered (and therefore has to be added to the table of contents semi-manually).

For code, the {\tt verbatim} environment (maybe in conjunction with $\backslash${\tt include}) is helpful:

\begin{verbatim}
It reproduces text
  exactly
    as
      it
        is
typed.
\end{verbatim}

Good luck and enjoy writing!