### 2009-11-11 docs/marcello-dissertation/thesis/source-code.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\chapter*{Appendix A: Triangulations of Polytopes}
%\addcontentsline{toc}{chapter}{Appendix A: Triangulations of Polytopes}

\appendix{Appendix A: DSP Data Persistence OSGi Bundle}

This section includes the source-code for the DSP Data Persistence. Each
section details one single main artifact used on the development of the
component. However, only the main ones will be added into this document, while
the remainder can be downloaded directly from the Subversion repository.

The source code is located in the Subversion repository for the NetBEAMS
project. This is how it is structured:

%    DSP_HOME - http://code.google.com/p/netbeams/source/browse/ main
%    repository for evaluation.
%    - $PERSISTENCE = DSP_HOME/browse/branches/marcello/persistence/ = defines
%    the project home for the persistence component.
%    
%    - PERSISTENCE/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence =
%    defines the directory for the persistence bundle.

\section{OSGi Bundle Build}

This artifact defines the ANT build file.

\lstinputlisting[language=Ant,label=persistence-builder,caption=Build
system using Apache
ANT]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/build.xml}

\section{DSP Data Persistence Service}

\lstinputlisting[language=Java,label=persistence-component,caption=DSP
Data Persistence
Component]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPDataPersistence.java}