### 2009-11-14 docs/marcello-dissertation/thesis/chapter9-appendix.tex

\chapter{Appendix - Source Code / Artifacts}

This section includes the source-code for the DSP Data Persistence. Each
section details one single main artifact used on the development of the
component. However, only the main ones will be added into this document, while
the remainder can be downloaded directly from the Subversion repository.

The source code is located in the Subversion repository for the NetBEAMS
project. This is how it is structured:

\section{DSP Data Persistence Component}

This artifact defines the ANT build file.

\lstinputlisting[language=Ant,label=file:dsp-build.xml,caption=Build
system using Apache ANT]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/build.xml}

%\lstinputlisting[language=bash,label=file:osgi-manifest,caption=OSGi
%Manifest
%Descriptor]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/META-INF/MANIFEST.MF}

\lstinputlisting[language=XML,label=file:dsp-config.xml,caption=DSP
Deployment
Configuration]{../../../../netbeams/versions/v2/config/deployment/config.xml}

\lstinputlisting[language=XML,label=file:dsp-matcher-config.xml,caption=DSP
Matching Rules]{../../../../netbeams/versions/v2/config/deployment/matcher_config.xml}

\lstinputlisting[language=XML,label=file:dsp-matcher-config-gumstix.xml,caption=DSP
Matching Rules
for
Gumstix]{../../../../netbeams/versions/v2/config/deployment/matcher_config-gumstix.xml}

%TODO: add bootstrap message in case it is implemented

\lstinputlisting[language=Java,label=file:dsp-activator,caption=OSGi
Activator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/osgi/DataPersistenceActivator.java}

\lstinputlisting[language=Java,label=file:dsp-component-service,caption=OSGi
Service - Component]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPDataPersistence.java}

\lstinputlisting[language=Java,label=file:dsp-mongo-service,caption=DSP Data to
Mongo DB
Service]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMongoCRUDService.java}

\lstset{language=XML,morecomment=[s]{!--}{--},label=file:dsp-message-serialized-ysi,caption=DSP
Message with a YSI sonde data payload}
\begin{lstlisting}
<?xml version="1.0" encoding="UTF-8" standalone="yes"?>
<MessagesContainer uudi="24929c29-60ee-4d17-af08-64d9446277ef"
        creationTime="2009-03-06T15:17:18-0800" destinationHost="192.168.0.106">
   <MeasureMessage ContentType="org.netbeams.dsp.ysi"
         messageID="435a61f6-370f-458d-aeb7-6e92270a79cb">
      <Header>
         <CreationTime>1236381438480</CreationTime>

         <Producer>   
            <ComponentType>
                 org.netbeams.dsp.platform.management.component.ComponentManager
            </ComponentType>

            <ComponentLocator>
               <ComponentNodeId>1234</ComponentNodeId>
               <NodeAddress>192.168.0.103</NodeAddress>
            </ComponentLocator>
         </Producer>

         <Consumer>
            <ComponentType>
                 org.netbeams.dsp.wiretransport.client
            </ComponentType>

            <ComponentLocator>
               <NodeAddress>LOCAL</NodeAddress>
            </ComponentLocator>
         </Consumer>
      </Header>
      <Body>
         <SondeDataContainer>
            <soundeData date="15:17:18" time="03-06-2009">  
               <Temp>21.20</Temp>
               <SpCond>193</SpCond>
               <Cond>179</Cond>
               <Resist>5588.40</Resist>
               <Sal>0.09</Sal>
               <Press>0.084</Press>
               <Depth>0.059</Depth>
               <pH>7.98</pH>
               <phmV>-79.6</phmV>
               <ODOSat>99.5</ODOSat>
               <ODOConc>8.83</ODOConc>
               <Turbid>0.4</Turbid>
               <Battery>8.7</Battery>
            </soundeData>
         </SondeDataContainer>
      </Body>
   </MeasureMessage>
</MessagesContainer>
\end{lstlisting}

\section{Experiment Implementation}

\lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Experiment
Implementation]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}

\lstinputlisting[language=Java,label=file:experiment-dsp-implementation,caption=Experiment
Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}