### 2009-11-16 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Conclusions and Future Works}

This work presented a solution to store data from sensor networks.....
...the main result of this work was that the persistence is

There are many important features covered in this work, including...
  - Use of open-source UI for KVP Sensor Data

Other important features that COULD HAVE BEEN covered, includes\ldots
  - Different Dababase system, use of cloud computing, etc
  - Use of MapReduce technology and implementations, distributed Processing that scales the proposed KVP storage

... The main results from this dissertation were important because of the contributions on new technologies...
  - An alternative Data Model for Data Storage in Sensor Networks
  - An alternative apprach for Data-Centric Storage in Sensor Networks

\section{TODO list}