### 2009-11-22 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Review of chapter 1 after Prof. Puder's feedback.
 - More abstract introduction, without mentioning keywords
   used in the work such as NetBEAMS, DSP, KVP, etc;

 - Reduced structure

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@655 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Introduction}

Sensor networks are commonly used in different areas for different purposes
such as scientific measurements and general population use \cite{sn-intro01}
\cite{sn-intro02}. In the scientific community, they serve as tools to monitor
the state of the environment by storing samples of data in specific periods of
time. For instance, the NASA Jet Propulsion Laboratory uses the Volcano
SensorWeb \cite{sn-ex02}, a sensor network which aims to collect data from
specific volcanoes in Alaska and South Pacific to be used with heuristics
regarding dangerous activities. Similarly, the National Data Buoy Center, a
department of the National Oceanic and Atmospheric Administration (NOAA)
\cite{sn-ex03}, provides online data from different buoys in different coast
shores around the world regarding water quality, tide information, etc, what
characterizes sensor network for environmental monitoring \cite{sn-ex01}.
Therefore, sensor networks can be seen as a very important scientific research
and commercial field with direct application and use of its produced data.

In order to provide access to the collected sensor data, different approaches
can be taken to interrogate sensor devices indirectly by accessing its data in
places called network sink. Similarly, the data model and database system
location may also influence in the process of reusing the collected data. In
this way, the lack of a persistence layer for existing sensor networks requires
a deep understanding about its infrastructure, the nature and location of the
collected data, the chosen data model representation, among others. For
instance, how can a data persistence layer safely represent the collected sensor
devices' properties without the need of restructuring any existing data model?
Should the end users' skills in database system be taken into account or the
use of programming languages to access the data \cite{sn-programming-language}?
This problem is currently faced by the departent of Biology at San
Francisco State University with a marine sensor network managed by
researchers without Computer Science skills. While the data access may impose
the difficulties, technical issues may also rise while the execution of a
sensor network such as growth in terms of number of sensor devices and,
consequently, the number of data produced. Last, but not least, the means to
access the data for reuse is the most important features of the sensor
network, taking into account the different data formats used by the main users
of the system. How can specific collected sensor data be exported to other
formats such as spreadsheets or archived files meant to be shared among
research peers? For these, and many other reasons, the inception of a
persistence storage can be a very challenging problem to solve.

In view of the fact that the persistence storage of a sensor network imposes
many different challenges, this work proposes the review of the current state of
the art for sensor networks in order to provide a data persistence layer for an
existing sensor network. As a matter of fact, this work's first contribution is
the inception of a set of basic taxonomies for data persistence in sensor
networks, proposed to drive an empirical evaluation of different technologies
used in the sensor network community. Furthermore, it also takes a look at
different alternatives not yet explored through experiments for a selected
sensor network as a case study. Similarly, this work makes another novice
contribution by suggesting a ``lighter" data model and database infrastructure
that easily captures sensor network device's observations, taking into account
the non-predictable scalability of its infrastructure, as well as the aspects of
the data's nature such as identity, time and location of the collected data.

This dissertation is organized as follows: the next chapter describes the
literature review as the state of art on sensor networks and data persistence, 
providing the needed foundation for the design of taxonomies in data
persistence for sensor networks in chapter 3. Then, chapter 4 examines the
selected case study and its requirements for a persistence layer, whose
technology properties are evaluated in chapter 5 against the developed
taxonomies. The culmination of this work is chapter 6, which details the design
and implementation of a data persistence solution for the selected case
study. Likewise, chapter 7 discusses the solution by presenting the
measurements and conducted experiments, and describing the findings
and contributions of this work. Last, but not least, chapter 8 presents the
conclusions of this work, together with the possible future works not addressed
here.