### 2009-11-27 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib