### 2009-11-29 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Revision of chapter 4.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@663 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Data Persistence for Sensor Networks: A Case Study for NetBEAMS}
\label{chap:netbeams-overview}

This chapter presents the primary motivation and scope of this paper. The
infrastructure of SF-BEAMS is presented as a real-world environmental sensor
network, deployed by the department of Biology at the San Francisco State
University. It serves as the foundation of the execution of NetBEAMS, a joint
venture between the department of Biology and the Computer Science. As an
alternative approach for data collection, NetBEAMS is used to automate the data
management and collection process of SF-BEAMS. However, it offered different
open problems, including Data Persistence. Therefore, the realm of the solution
presented in the following sections and chapters are based in the case study
presented in this current chapter.

\section{SF-BEAMS: a Marine Sensor Network for Water Quality Monitoring}

The San Francisco Bay Environmental Assessment and Monitoring Station, or
SFBEAMS \cite{sfbeams2006}, is a marine sensor network operated by the Romberg Tiburon
Center (RTC), whose primary focus is the study of complex marine and estuarine
environments using the SF-BEAMS sensors. The network is deployed offshore of
its pier located on the banks of the San Francisco Bay at the Tiburon Island,
California, being operated by the department of Biology from San Francisco
State University.

\subsection{The SF-BEAMS Infrastructure}

The SF-BEAMS sensor network is responsible for providing data for water quality
monitoring, as well as weather and surface conditions. Figure
\ref{fig:sf-beams} shows a picture taken from the SF-BEAMS web-camera.

\begin{figure}[!b]
  \centering
    \includegraphics[scale=0.7]{../diagrams/cam_image-oct15}
  \caption{Picture of the SF-BEAMS Network, at the RTC pier. Tiburon, CA.}
  \label{fig:sf-beams}
\end{figure}

In general, the SF-BEAMS network infrastructure contains varying wired and
wireless devices standing on pylons docked at the pier. Each of them is
responsible for observing different conditions of the area, having its own
mechanisms for internal storage for the collected data. In this way, data can
be directly transferred to the labs via Ethernet cables or collected manually
by a laptop computer.

\subsection{SF-BEAMS produced data and collection process}
\label{sec:sfbeams}

As described in section \ref{sec:sn-infrastructure}, sensor devices produce
its observed data based on properties of measurements defined by its
manufacturer. SF-BEAMS can be seen a single-hop star sensor network containing
different instances of sensors, including the ``YSI 6600 ESD V2", as seen in
figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
device that produces around 52 bytes (13x4 Bytes) on a single real-time data
stream as shown on table \ref{tab:ysi-data-stream}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.7]{../diagrams/ysi-device}
  \caption{Picture of the case study sonde device: YSI 6600 ESD V2}
  \label{fig:ysi-device}
\end{figure}

\begin{table}
    \caption{YSI Data Stream}
    \begin{center}
        \begin{tabular}{lr}
  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83 
  0.4     8.7
        \end{tabular}
    \end{center}
    \label{tab:ysi-data-stream}
\end{table}

In order to have access to the observed data, the network staff downloads the
data using one of the device's connection such as the RS-232 serial connector
\cite{rs232}. Then, the RTC labs receives the collected data transferred from
the sensor devices to be used for archival. The data is described, indexed and
distributed to its main users. According to one of its staff members, the
infrastructure of the SF-BEAMS in May 2009 was comprised of the following:

\begin{itemize}
  \item Among other devices, 5 YSI sondes in operation at the RTC pier;
  \item Each YSI sonde produces 52 Bytes for each its observations;
  \item The sampling frequency rate could be configured in ranges of 1, 6 or
  15 minutes, depending on the specification;
  \item Data for final users is published with Provenance Data regarding time,
  without description of the origin of data.
\end{itemize}

Considering the infrastructure previously described, the size of the data
produced by the YSI sonde sensors at the RTC pier can be estimated by its
measurement frequency as follows:

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Amount of data produced by the RTC's YSI sondes}
        \begin{center}
        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
        \textbf{Weeky} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
        \end{tabular}
        \end{center}
\end{table}

As it is shown, the data produced by this type of sensor runs in the order of
at most hundreds of Megabytes a year for a maximun of \textbf{483,840}
samples, what can be considered a exceptionally sparse data storage
requirement as compared to other networks. Upon collecting data from sensors,
the RTC staff use automation scripts written in Matlab to process, index and
distribute the raw data in different formats. One of these formats is the
OPeNDAP \cite{opendap}, a format widely used at research institutions that
promote easier data exchange among them. The access to the archived data is
done by using the OPeNDAP server is located at
http://sfbeams.sfsu.edu:8080/opendap, as a request and the corresponding
output for one of the YSI devices can be seen in Listing \ref{file:rtc-ysi-opendap}.

\subsection{Taxonomic Classification of SF-BEAMS}

In order to distinct each of the properties of the SF-BEAMS sensor networks,
consider the following list as a classification of the data persistence
mechanisms from the sensor network used by the project:

\begin{itemize}
  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is solely
  used for Data Archival purposes, since they are stored in the RTC's lab as
  part of the collaboration research project that monitors each part of the
  California coast. The data sets are distributed by a specific date in an
  annual calendar;
  \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
  of data is directly related to its location and, as a consequence, the
  strategy of \textbf{External Data Storage} is used to store its collected
  data at the network sink at the RTC's labs;
  \item \textbf{Data Model}: no data model is used by the RTC's staff to model
  the data. A straightforward conversion of the data collected from the sensors
  directly to spreadsheets, using comma-delimited values, is used. That is, the
  data mode used \textbf{Schema-less};
  \item \textbf{Data Provenance}: data is stored in files with conventions that
  take the time when the data was collected. The contents of the files contain
  the \textbf{Data Identity} and \textbf{Time Dimensions}, as shown in the
  samples collected and shown in Listing \ref{file:rtc-ysi-opendap};
  \item \textbf{Query Processing Mechanism}: data is queried in a
  \textbf{Centralized} fashion, due to the nature of its purpose and locality;
  \item \textbf{Database System Organization}: there are no databases in use.
  The SF-BEAMS staff use the OPeNDAP Hyrax server as a middleware used to read
  and interpret the collected data. 
\end{itemize}

\section{NetBEAMS: a component-based approach for SF-BEAMS}
\label{sec:problem-requirements}

Although the execution of the RTC's sensor network can be operated as it
described in the previous sections, \cite{netbeams2009} described some
operational challenges faced by the RTC staff during regular activities during
the data collection process. It was clear to the research group that SF-BEAMS
could have not only its data gathering process automated, but also its data
management and distribution. By the use of COTS\footnote{Common-Off-The-Shelf
designates a product that is produced and sold in bulk} embedded devices and
open-source \cite{open-source} software, the research group developed a second
version of NetBEAMS, a component-based approach suggested improving
operational activities from SF-BEAMS using systems automation. In brief, one
of the open problems of NetBEAMS was regarding Data Persistence, the primary
motivation of this dissertation.

The Networked Bay Environmental Assessment and Monitoring System, or Net-BEAMS,
is a work-in-progress joint project proposed by the department of Computer
Science and the RTC at San Francisco State University. The main reason for the
inception of NetBEAMS is the potential to reduce operational costs and
facilitate the data gathering process. In the current state of SF-BEAMS,
Biologists need to obtain data manually from the sensor devices in order to
take them to the network sink for further processing. For this reason, the Data
Sensor Platform (DSP) \cite{netbeams2009} was proposed as the system
architecture that can address the operation of SF-BEAMS. In this way, in-depth
documentation describing the DSP Platform, including its architecture and data
gathering process, is provided in the appendix section \ref{sec:dsp-details}.

The scope of this work then is defined as to provide a persistence layer for
NetBEAMS, making a selection of a persistence layer that is a exception to the
taxonomical classification of SF-BEAMS. Furthermore, the technology to be
selected for the persistence layer must take into account the requirements of
NetBEAMS primary users such as Computer Science researchers, as well as end
users such as Biologists from the RTC laboratories.

\section{Requirements for NetBEAMS Data Use}

The main reason for the operation of SF-BEAMS using NetBEAMS's automated
approach is that it can potentially reduce operational costs and improve the
data collection process. In review of this, the scope of a persistence layer
can be summarized as a group of functional and non-functional requirements in
order to select a database system for the sensor network supported by NetBEAMS.

\subsection{Functional Requirements}

\begin{itemize}
  \item Reuse the NetBEAMS infrastructure and develop a component responsible
  for data persistence in a database system;
  \item The Persistence System must obey to the characteristics of the
  categorization of SF-BEAMS used by the taxonomies.
\end{itemize}

\subsection{Non-Functional Requirements}

\begin{itemize}
  \item The data model used to describe the data must not impose restrictions
  to the users of the system: Biologists and students without expertise in
  Database Systems;
  \item Data Representation must be similar to the ones used by RTC, with a
  more human-readable format;
  \item Scalable Systema low degree of maintenance, in a way that data
  gathering service should be able to manage the system with less service interruption;
  \item Data must be searchable in near-real-time with outstanding performance,
  as compared to the time when data was collected to the time when they are available;
  \item Easy data integration and export capabilities must be also taken into
  account;
  \item The database system must be free of charge, following the
  implementation specifications of NetBEAMS.
\end{itemize}