### 2009-12-01 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Experimental Results: Correct Behavior and Performance}

In order to evaluate the mongoDB, the different taxonomies identified are
verified from the logs and experiment setup. The implementation of the Data
Persistence Component and the mongoDB infrastructure are described in section
\ref{sec:dsp-data-persistence-implementation}.

The purpose of Sensor Data is used for Data Archival, since the data from the
sensor data categorized by the date and time it was collected. The location
of the collected data from the sensor devices are stored using the External
Storage, or using the Data-Centric Strategy. In this way, one of the advantages
of the denormalized data model provided by mongoDB is that the data can be
distributed using Horizontal Partitioning.

\section{Experiment Setup}

Experiments were conducted to evaluate mongoDB against the taxonomies defined in
chapter 3 by creating a general shell script
\ref{file:experiment-setup-executor} that sets up the environment by
initializing the mongoDB server and indirectly inserts the a desired number of
randomly generated YSI Data types \ref{file:random-ysi-data-generator}
implemented in Java as part of the NetBEAMS data collection mechanism as shown
in section \ref{sec:dsp-payload-implementation}. First, the definition of a 
document will be conducted according to the definition of the Data Provenance
definition of chapter 3. Then, the infrastructure of the database will be
evaluated using different scenarios.

\subsection{Key-Value definition}
\label{sec:keys-definition}

In document-oriented databases, there is no definition of table. In constrast,
the entities are persisted using any data arrangment, in any order, attributes
and the associated values, as shown in Listing
\ref{file:mongodb-ysi-data-format}. The definition of the document for mongoDB
takes into account the pairs of key and values related for the collected data
from the sensors, as well as the properties defined by the data provenance
study \cite{sn-provenance}. The collected data is stored in a key called
``observation''. On the other hand, the provenance-aware keys are defined as
follows:

\begin{itemize}
  \item \textbf{What: Data Identity}: the identification of the data captured
  from the sensor, as well as the DSP Platform message. The keys are comprised
  by the attributes ``\underline{ }id'' as an unique identifier provided by
  mongodb, the ``message\underline{ }id'';
  \item \textbf{When: Time Dimension}: the time dimensions give information
  regarding when the data was collected and transferred to the network sink.
  They are the ``time.valid'', ``time.transaction'';
  \item \textbf{Where: Data Location}: the location where 
 the data originated. The ``sensor.ip\underline{ }address'',
 ``sensor.location.latitude'', ``sensor.location.longitude'' give information
 regarding the .
\end{itemize}

\subsection{Workload}
\label{sec:workload}

The workload selected for the experiments reflacts the current use of the data
at the SF-BEAMS infrastructure. A number of data worth a year of collected data
from one single YSI device is randomly generated and inserted at the database.
Then, five additional run will be performed and investigate different search
scenarios for performance.

\begin{itemize}
  \item \textbf{First Round}: 483,840 items worth one year of produced YSI data
  at the rate of 1 observation per minute;
  \item \textbf{Next Rounds}: 483,840 * 5, or 2,419,200, representing 5
  different YSI components producing data every minute.
\end{itemize}

\subsection{Single Server Environment}

The definition for the single server follows a regular specification of the
mongoDB database system. All the collected data from sensors are stored in a
given single server, which characterizes an External Storage device.

\subsection{Data-Centric Environment}

The definition of a Data-Centric deployment follows the specification of the
mongoDB shard servers, where in contrast to the single server, all the
collected data is spread out to different mongoDB clients that manage
partitions of servers on a cluster.

As described in the taxonomy reviews, the definition of the sharded system must
include the definition of a shard key, which is responsible for partitioning
the instances. In this way, the shard key is defined as the value of the
attribute ``sensor.id\underline{ }address''.

\subsection{Scenarios}
\label{sec:exp-scenarios}

In order to verify the feasibility of the system, the use of different search
scenarios are used defined based on the use cases of section
\ref{sec:use-cases}:

\begin{itemize}
  \item A) Find data by specific date ranges on either the valid or the
  transaction time;
  \item B) Find data by specific IP address, as it identifies a given sensor
  device and any data produced;
  \item C) Find data by specific values of the observed attributes of the sensor
  device such as temperature, salinity, etc.
\end{itemize}

Moreover, the experiments must also export data in different formats such as 
spreadsheets and use different programming languages APIs to access the data.

\section{Measurements Results}
\label{sec:exp-measurements}

The result of the insertion of data in different scenarios are shown in the
different logs captured during the experiments. First, the avarage insertion
times gives the capacity of the system with and without the definition of
indexes over the data. Then, the numbers regarding the scenarios described
earlier will be shown.

\subsection{Single Server Environment}

The insersion time avarages an impressive ~169,508 instances per minute during
the experiments using the Java driver to connect with the mongoDB server. The
following lists the storage consumed by the generated data described in the
section \ref{sec:workload}.

\begin{itemize}
  \item 1 YSI worth of data = 234.16 MB 
  \item 5 YSI worth of data = 1.23 GB
\end{itemize}

On the other hand, the avarages for each of the scenarios are devided by the
amount of data produced by one single device or five different devices. For
data worth five different devices, without indexes in the database, the values
can be summarized as follows:

\begin{itemize}
  \item A) = ``time.valid'' for week period, or 7 days = 3820ms
  \item B) = a given IP address by ``sensor.ip\underline{ }address'' = 3721ms
  \item C) = given values for the keys ``observation.Salinity'' and 
  ``observation.WaterTemperature'' = 3167ms
\end{itemize}

\section{Discussion}

The implementation of a data persistence for any sensor network must take into
the characteristics of the sensor network, the location of the collected data,
etc. As described in chapter 4, the taxonomies classify different aspects of
the persistence of data. Along with it, the design and implementation of a data
persistence layer for NetBEAMS must meet the requirements and specifications
described in chapter 5 and chapter 6, respectively. This section discusses the
experiments scenarios described to evaluate a real-world experiment relating
the implementation and results to each of the taxonomies. Furthermore, the
experiments provided by this work cover a one year worth of data for one and
five YSI devices and evaluates the use cases defined in section
\ref{sec:exp-scenarios}.

\subsection{Taxonomic Evaluation}

Any database system covers the purpose of the sensor data for NetBEAMS, which
is Data Archival, along with the type of storage for the data as an External
Storage. After selecting mongoDB, the development of the persistence layer met
the requirements of archiving data after the randomly generated data for the
YSI sonde was persisted. As it is shown in section \ref{sec:exp-measurements},
the collected data is archived in files using the BSON format \cite{bson}, the
binary representation used to store instances of the document described in
Listing \ref{file:mongodb-ysi-data-format}. Along with the purpose of data
archival, the use of mongoDB also met the specification of an external storage
since the data persisted in the file system.

One of the requirements for NetBEAMS is the use of a schema-less data model.
mongoDB uses a variation of the key-value data representation based on the JSON
\cite{json}, and provides APIs in different languages for the data creation
process. In this way, the implementation of the persistence model captures the
instances of any known data collected from NetBEAMS and saves it using the
document pattern as shown in Listing \ref{file:mongodb-ysi-data-format}. One
application of a schema-less data model is the addition of any new key-value
pair to any instance of the document, as mongoDB supports update operations to
any of the collections. For instance, annotation in sensor networks is
widely used to describe the findings of a particular data set
\cite{sn-annotation-ex1, sn-annotation-ex2}. In case of an oil spill in the San
Francisco Bay occurs, researchers can update the data set of a given date range
by adding tags to the relating documents. As a consequence of the use of a
schema-less data model, these changes do not affect the other data instances,
since there are no schema changes.

As described in chapter 5, the schema-less data model can provide a better
support to Data Provenance because the data model gives support to the change
of records without affecting the others. In this way, the metadata used by the
implementation described in the previous chapter clearly describes the three
fundamental questions regarding the collected data from sensor devices, as
described in section \ref{sec:keys-definition}. First, the identity of the data
is given by the key ``message\underline{ }id'', in a way that it uniquely
identifies the data produced by the sensor identified by the key
``sensor.ip\underline{ }address''. In this way, the message can be tracked back
to the sensor that produced the data for different purposes such as tracking
the data producer. Second, time dimensions were used in order to track two
different aspects of the data collection: the valid time can be used to
correlate observations to the exact moment it occurred, while the transaction
time can be used to verify when the data was collected, and make decisions such
as deleting the instances of data were collected in the previous two days, etc.
Last, but not least, the use of the key ``sensor.location'' can give the exact
location of the data, since researchers can be interested in finding data based
on a specific location. Finally, the most important aspect of the schema-less
data model is that the observed data is added to the key ``observation'', as
any sensor device carries its set of attributes of key-value pairs. A good
example of the applicability of the schema-less data model is the scalability
regarding changes to the data structure. For instance, if an existing sensor
device is flushed with a new version of the firmware\footnote{The first set of
machine instructions to run on the hardware after the application of power stored
permanently in PROM or ROM or semi-permanently in EPROM}, changing the format
of data types or adding new values, the following instances of collected data
may be different to the previous one by simply adding the new additions of
data. For this reason, the DSP Data Persistence gives administrators the
ability to choose which fields to persist by declaring it using a bootstrap
message for the component, as shown in the parameter
``YSI\underline{ }DESIRED\underline{ }PROPERTIES'' of Listing
\ref{file:dsp-data-persistence-bootstrap.xml}.

The centralized query processing used by NetBEAMS is a result of the
infrastructure used by NetBEAMS, as well as the purpose of data and the
location of the collected data. As a direct result of a schema-less data
model, the centralized query processing is easily developed using any of the
available programming languages provided by mongoDB. Considering the fact that
programming languages are easier abstractions to researchers without computer
science background \cite{sn-programming-language}, mongoDB's approach to data
access through API can be seen as an easier task to manage data. As it is shown
in Listing \ref{file:experiment-query-scenarios}, the scenarios designed to
evaluate the implementation are written in Javascript, since it is the script
language by the mongoDB Shell. Therefore, it attends the requirement of not
using the SQL language as the main query language of the data.

One of the most important features of a data persistence for sensor
networks is the replacement of the SQL by the use of programming
languages, as they tend to be easier abstraction to users without the
background in database systems \cite{sn-programming-language}. In this way,
mongoDB covers the requirements of providing data access to the collected
sensor data in different languages. For example, the data collected by the DSP
Data Persistence component is written in Java, while the implementation of the
experiments scenarios for querying the collected data is given in Javascript. In
this way, researchers of the RTC may choose one of the enumarated languages
APIs provided by the mongoDB to access the stored data from NetBEAMS.

Finally, the implementation supported by mongoDB is either in a single or
distributed servers. The former approach is done as a regular database server
on a single host machine, where its main goal is to manage the collected sensor
data. On the other hand, however, the latter approach can be used to improve
the performance of the data selection in larger data sets by offering different
capabilities such as Database Shards and Data Replication.

\subsection{General Evaluation}

In general, the implementation of the DSP Data Persistence component meets
the specification of the use cases defined in section \ref{sec:exp-scenarios}.
The implementation of mongoDB using a single server environment can give
support to the current workload and infrastructure from the use of NetBEAMS to
collect data from the sensors. From the perspective of disk-space utilization
and given that ~1.5 Gigabyte of disk storage is necessary to store
one-year-worth of collected data for five YSI devices, the implementation
scales to 80 years of data in a server with 120 Gb of available disk space.
However, if the database server runs out of space, the addition of a new
database shard can be done.

The performance to insert data into the database system using the Java driver
is documented by mongoDB as the fastest approach, providing persistence to
approximaly 169,508 instances of data. Considering this number of insertions to
the database, it is clear to say the mongoDB implementation can virtually
support around 169 thousands of sensors a minute. Similarly, the implementation
of a data-centric scenario using mongoDB can improve the performance of the
system, given that the schema-less data model gives the ability to partition
data horizontaly. In this way, latency related to data constraints and data
integrity checks are avoid because an entire instance of the collected data
resides in the same ``chunck'' of data. This is the main advantage over the
relational model, previously discussed in chapter 5.

One possibility of improving performance on the given implementation is the use
of indexes on each of the keys. As a direct consequence of this procedure,
insertion time to this approach is approximaly decreased by a 10th of the
insertion time experienced with the experiments without indexes. Similarly, it
turns out that the use of indexes improves the query processing by 5 times when
indexes are not used. Therefore, the use of indexes should be enforced in a
real-world deployment, since both requirements for insertion and retrieval are
met.

When it comes to re-using the data collected by the DSP Data Persistence,
different approaches are available. First, the use of the export capabilities
to a common-delimited values was performed \ref{file:mongodb-export-command},
taking approximaly 3.3 minutes to export one year worth of data for 5 YSI
devices. Second, the use of the client shell gives a centralized access to
either the single server of sharded server option. However, in order to better
provide access of data to users without skills in database systems, web
applications can be built on top of the REST \cite{http-rest} Web Services APIs
provided by mongoDB. One such example of using the REST API is the adaptation
of Apache Futon 4 mongoDB, as shown in Figure
\ref{fig:view-collections-instance-browser-futondb}, where users can navigate
through the collections of the collected sensor network data categorized by the
collection. Therefore, RTC staff technicians could take advantage of this tool
to view the collected data.

The design and implementation of the guidelines designed by the taxonomies can
be applied to any sensor network that produces data with the purpose of Data
Archival in External or Data-Centric storage.