### 2009-12-02 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{State of the Art}

In general, this chapter reviews sensor networks characteristics and
properties, necessary to the understanding of the basic properties of a sensor
network. Moreover, it reviews the characteristics of data persistence for
sensor networks, characterizing the different approaches used to store data
such as data model used and how the collected data is described.

\section{Survey in Sensor Networks and Its Data}

Sensor Networks are specialized types of network systems comprised of nodes
with specific goals such as to produce and collect data. The former is seen
as a sensor device capable of producing data to the network based on an
observation from the environment or just a raw value regarding a given
propertiy, while the latter can be seen as the network sink responsible to
receive the observed or ``sensed'' data. In this way, data persistence in sensor
networks depends on the approaches used to store the data on the data sink,
since these types of nodes can be located in the network close to the devices 
or in a centralized location. As a consequence, the query used to query the
data can take place in the network nodes themselves or in the centralized node.
The former approach is used when the data is collected at the sensor node on
its local storage \cite{sn-storage01, sn-storage03}, strategy typically
used to maximize the sensor's resources, while minimizing energy consumption
\cite{sn-storage04}.The latter strategy, however, has been used when the
collected data is archived in a centralized database system for later reuse
\cite{sn-storage02}.

When it comes to data representation, the relational data model
\cite{relational-model} is one of the most used approaches to represent data in
sensor network databases in any of the query processing strategies and, in most
cases, with a modified version of the SQL language \cite{sn-db-newop}. However,
the design of the database system must be done with prior knowledge of the
sensor types when choosing the relational data model, as the database design
must be normalized \cite{db-normalization} for the entities already identified
for the system. For example, the introduction of a new sensor device might
represent a potential change on the database structure by seeking a new
normalized version of the database design. In addition to challenges of
maintaining the database schema normalized, the use of Data Provenance 
approaches in sensor network data usually addresses the problem related with
the origin of data and its entire history \cite{sn-provenance}.

The following subsections describe the main characteristics of sensor networks
described in the literature, and then others that interest data persistence in
sensor networks.

\subsection{Deployment and Mobility}

Each node of the sensor network can be deployed in an arbitrary method such as
being dropped from an airplane, or in a manually placed at chosen geographic
locations \cite{sn-intro01}. Moreover, the deployment can be done as a one-time
event or as an iterative process, depending on the needs of where the sensor
device needs to be placed.

Nodes can be able to move around the deployed locality or remain fixed at a
single place. In the case of a sensor networks comprised with different moving
sensors, a given sensor node can move occasionally or continuously, as a result
of an accidental side-effect or a desired property of the device. Usually,
these sensors are automotive or they can be attached to another moving object.
For this reason, \cite{sn-intro01} describes those types of mobility as active
or passive.

\subsection{Device Size, Resources, Cost, Energy, Heterogeneity}

Such sensor nodes can be of different sizes, ranging from microscopically small
particles, small devices to bigger ones with a few feet cubic. In addition to
being limited in size, they are also limited in resources and computational
power, as they are designed to either stored energy on batteries or
use energy scavenged from solar cells \cite{sn-intro01}.

Sensor devices range in size, capabilities and other properties, which are
associated to their cost ranging from hundreds to thousands of dollars. All in
all, the price depends on the device's functionality, energy utilization
capabilities, among others. For this reason, \cite{sn-intro01} explains that a
sensor network may potentially contain similar sensor devices, as they can
contain dissimilar ones from each other, which characterizes the heterogeneity
of a sensor network by the hardware capabilities such as GPS enabled or having
GSM network connectivity.

\subsection{Communication Mode, Network Infrastructure, Coverage and
Connectivity}
\label{sec:sn-infrastructure}

Sensor nodes can communicate in different modes such as radio, diffuse
light, inductive or sound, which may influence on the type of protocols used to
communicate among the nodes, where they can be said to be in an organized
infrastructure or as an ad hoc way \cite{sn-intro01}.

When it comes to infrastructure, sensor networks differ in topology:
single-hop, star, networked stars, tree and graph are examples of these
topologies. Each of they topologies may influence on how data is collection
and routed to their data sink. While there are only two levels of node
connectivity in sensor networks deployed as single-hop and single start
topologies, networks whose topology is either tree, graph or interconnected
stars use communications among each of their nodes for data communication.

Last, but not least, the coverage of a sensor network is related to the
physical space that is ``visible" to the sensors, which is directly related to
the purpose of the network, its size in terms of the number of nodes that are
deployed and the sensor network lifetime. In this way, sensor devices can be
deployed in different areas and distributed in sparse or dense manners, where
the sensor network they support is providing the necessary data that
represents the coverage of the network. However, considering that some types of
sensor devices are very likely to regular hardware failures, projects have
provided redundant nodes \cite{sn-intro01} to provide better data coverage.

No matter the type of coverage provided by a sensor network, different levels
of coverage may also influence on the connectivity of the sensor nodes. In
order to provide connectivity to redundant nodes, for example, a physical 
connection must be available at all times. On the other hand, sensor nodes can
also communicate with other nodes in a sporadic way when they are designed to
be isolated with only communication activities to report the collected data to
their immediate peer (other nodes and/or data sink).

\subsection{Network Size, Lifetime and Quality-Of-Service}

The network size is related to the degree of coverage of a sensor network, and
it is measured by the number of devices deployed as nodes of the network.
\cite{sn-intro01} reports examples of sensor networks with sensor nodes ranging
from dozens, hundreds or even thousand of network nodes.

The application of the sensor network determines for how long the sensor network
may be exist. For example, the existence of an environmental sensor networks
\cite{sn-ex01} may be not determined since its the data collected from the
network are used for archival purposes, hence, without termination of the
execution of its sensing tasks. On the other hand, however, the lifetime of a
sensor network designed to count the number of people who entered in different
rooms of a building during an event has its life-time related to the period of
the event.

Finally, the design of sensor networks can be related to the constraints
related to quality-of-service (QoS) such as real-time delivery within a
certain period of time or robustness such as maintaining data alive even
though there a communication link is not available \cite{sn-intro02}.

\section{Persistence Storage for Sensor Networks}

The most valuable objective of a sensor network is related to its first
product: the use of the raw data. Data collection in sensor networks is
directly dependent on the specifications of the sensor network, its routing
mechanisms and the final destination of the produced data. In a nutshell,
\cite{sn-storage03} describes data as follows:

\begin{itemize}
  \item \textbf{low-level property} as the raw data collected from the sensor
  device; For example, the sea level as a number;
  \item \textbf{Observation or Event} as the processed data in the format of
  the final observation. For instance, if the ocean is in its highlight or low
  tight.
\end{itemize}

\subsection{How the Collected Data is Used}
\label{sec:sn-data-purpose}

At first glance, the raw data serves as the main outcome of the sensor network.
However, the first consideration that must be considered is regarding the
purpose of the collected data, as \cite{sn-provenance} explains differentiates
between two main benefit of the sensor data:

\begin{itemize}
  \item \textbf{Real-time Data Stream}, as the generated data from the sensor
  is not transmitted, but extracted from the device as it is requested. The
  data is still in its standard form specified by the sensor manufacturer;
  \item \textbf{Archival Data}, as the collected data include historical
  significance or will be use by other applications at a later time.
\end{itemize}

While real-time data stream are used and can potentially be discarded, the
archival data needs to be stored in a secondary memory. Nonetheless, given the
nature of the sensor device specification, the low-level data produced by a
given sensor device might not include vital information regarding the identity
of the data itself. The use of Data Provenance techniques is described by
\cite{sn-provenance} as an essential step to provide the collected data
specific metadata regarding its location and lineage.

Under these circumstances, the collected data needs first to face its journey
starting from its transmission from its producing device to the specified data
sink as specified by \cite{sn-storage01, sn-storage02, sn-storage03}.

\subsection{Storage Location for Collected Data}
\label{sec:sn-storage-locations}

Given that collected data can be used for different purposes as described in
section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
responsible to send the raw data to a specified data storage target. Upon
producing data, a sensor network node is specified to deal with the produced
data where, in most cases, depending on the sensor network infrastructure used
\cite{sn-storage03} (see section \ref{sec:sn-infrastructure}). In such a way,
different storage mechanisms are provided:

\begin{itemize}
  \item \textbf{External Storage}: the sensor device sends the produced data to
  another sensor node that contains an external storage.
  \item \textbf{Local Storage}: as it is the case of real-time data stream,
  where data is allocated on the sensor device's main read-only memory.
\end{itemize}

It is clear that the data flows from the sensor node to another depending on
the definition of the network. On the whole, the sensor network infrastructure
dictates where the data will flow to, as described in section
\ref{sec:sn-infrastructure}. \cite{sn-storage02} sees a wireless sensor
networks as a many-to-one data collection pattern, when its organization is
based on single-hop or star. On the other hand, when the data organization is
structured as a graph, tree, etc, the data flows from one intermediate sensor
node to another until it reaches its final storage node as described by
\cite{sn-storage01, sn-storage03}.

Last, but not least, yet a different storage strategy can be used: Data-Centric
Storage. This strategy aims at clustering the collected data based on a given
nature of the data such as location. Regardless the data storage mechanism
used, the query processing will depend on the sensor network infrastructure and
data location.

\subsection{Query Processing for Storage Locations}
\label{sec:query-process}

Given the different locations where the collected data might be as described in
section \ref{sec:sn-storage-locations}, different query processing for the
collected sensor data can be used as described by \cite{sn-storage03}:

\begin{itemize}
  \item \textbf{In-Network Query Processing}: sensor nodes that provide
  real-time data stream usually replies to requests to properties.
  Likewise, when the data is used for archival purpose, the data can be
  located at a network node specialized in storing the neighbors
  collected data. In this way, when a query is issued to the sensor
  network, the query processor will collectively verify every storage
  node for the collected data.
  \item \textbf{Centralized Query Processing}: a leading network sink receives
  the collected data from the peer nodes. In this case, the query processor
  queries the solo centralized node for the global values of the
  participating sensor devices.
\end{itemize}

There are pros and cons related to both of the approaches of query processing,
which are also related to the infrastructure, sensor network nature and
requirements. The primary purpose of the in-network query processing is to
minimize the number of data transmitted from the sensor devices to the final
data sink, as sensor devices are generally limited by constraints such as
energy utilization \cite{sn-storage03}. Along with this strategy, the literature
describes different proposals that aim at upgrading the Structured Query
Language (SQL) adding location-sensitive clauses to the language \cite{sn-db-newop}.
However, this crucial problem related to this method is that it may potentially
flood all nodes with query messages \cite{sn-storage04}.

In order to mitigate the problems related flooded nodes in in-network query
processing, the use of a Data-Centric storage strategy has been shown to be a
proven alternative for the purpose of energy consumption
\cite{sn-storage03, sn-storage01}.Its primary goal is to organize the
collected data based on a particular property, say location or event, in order
to be sent to individual storage nodes that relate to the observation. As the
main outcome of this strategy is to create network partitioning strategy, the
query processor would hit specific sensor nodes with the needed data and,
therefore, would decrease the load and maximize the capacity on the actual
sensor devices.

A centralized network query processing processes data in a network node
specified as the network sink. The network sink usually is a network node with
a database server with the collection of data delivered by all the sensor
devices from the network. As a consequence of this many-to-one communication
strategy \cite{sn-storage02}, the query process does not affect any of the
sensor nodes in the network and, thus, maximizes the sensor device
utilization. Moreover, the use of the data for historical purposes is easier
to manage and provide access to the primary users of the data, following the
process of data processing, compression, aggregation, etc
\cite{sn-db-modeling02}.

Although this approach seemed to be appropriate to the point of view of the
sensor nodes and used as the tradeoff between achieving capacity and energy
consumption, the centralized data collector may suffer from the unavoidable
development of a point of traffic concentration \cite{sn-storage02}, also
referred to as the phenomenon of congestion, or funneling effect
\cite{sn-storage04}. For this reason, simple distributed system techniques
such as the use of master-slave or data replication can be used to reduce the
overhead.

\section{Data Models and Query Engines for Sensor Networks}
\label{sec:data-models}

Different approaches are used to represent the collected data from sensors. The
use of basic data descriptors has been used in research groups which expertise
is not database systems, while others use such technology and the way to
represent data using more pragmatic approaches. \cite{sn-data-model-survey}
have shown different data models and the advantages and disadvantages. The 
following is a overview of these approaches used by the sensor network
community.

\begin{itemize}
  \item \textbf{Tabular Data Model}: comma-separated values are immensely
  popular due to its simplicity. Although data is not index , it is an easy
  approach for data exchange among researchers \cite{sn-provenance};
  \item \textbf{Relational Data Model}: it is used in the majority of the
  projects. Most of them due to the easy access to database systems, with
  refactored versions that support the implementation of sensor networks and
  its requirements \cite{sn-db-tinydb};
  \item \textbf{Structured Data Model}: similar to the relational model,
  structured data models are becoming popular due to the Internet. The use of
  XML data \cite{xml} is not yet popular, but there are a few projects using
  it \cite{sn-xml-usage01}\cite{sn-xml-usage02}, implementing its query by
  using a middleware \cite{sn-xml-middleware, sn-xml-query-engines}.
\end{itemize}

\subsection{Tabular Data Model and Query}

Simple use of Comma-Separated Values is an example of sensor network data. Data
is structured by the top header, and the values are in each line of the file.
The implementation of the SF-BEAMS \cite{sfbeams2006} uses this approach. In
summary, in a tabular data model

\begin{itemize}
  \item data is represented by columns in a flat file separated by a character
  that separates the values. One example of such data in sensor networks is the
  OPenDAP \cite{opendap} format, used by the SF-BEAMS project
  \cite{sfbeams2006};
  \item data is queried using a text editor file, without any indexing
  associated with it.
\end{itemize}

\subsection{Relational Data Model and its Query Mechanism}

Many different projects have used the relational model \cite{relational-model}
as the representation of data for sensor networks. In general, the following
properties apply to sensor networks:

\begin{itemize}
  \item data is structured in tables, and sensor properties are represented by
  the columns of these tables. Data instances are represented by tuples or rows
  of these tables. Finally, table relationships are implemented based on the
  relational algebra theory, which is the mechanism used to describe the
  relationship among different data types.
  \item data is queried using the Structured Query Languange (SQL) and uses the
  definitions of the tables and columns to extract data.
\end{itemize}

Before the creation of such a model, a database normalization process is
necessary to describe the data in a convenient way. An example of such use of
the relational model is the TinyDB database systems \cite{sn-db-tinydb}, which
uses the relational model to represent the types of sensors. This is a popular
choice because of the vast availability of research on relational database
models.

Although the implementation of a relational database seems to cover the needs
of data representation, \cite{sn-data-model-survey} described the problems
related to this approach. One such problem is that the relational model is not
suited for modeling time series data and metadata related to identify, time
and location. For problems related to it, came the implementation of different
approaches that add semantics to data types and different query clauses to the
Structured Query Language (SQL) to solve join operations \cite{sn-db-newop}.

\subsection{Structured Data Model and its Query Mechanism}

The use of XML is used to represent data in sensor networks has become popular
due the popularity of this data format. In general, the use of XML for
persistence can be summarized as follows:

\begin{itemize}
  \item data is structured in the collected data properties can be represented
  through the XML \cite{xml} instances by using tags, having its model validated
  by the use of XML Schema \cite{xml-schema}.
  \item in order to extract data from XML-based databases, the use of XML XPath
  \cite{xml-xpath} is used to query XML documents.
\end{itemize}

Some examples of the application of XML in sensor networks in
\cite{sn-xml-usage02} some have investigated the effective use of XML in
sensor networks to represent the collected data \cite{sn-xml-usage01}. An
example of querying XML data in sensor networks described by
\cite{sn-xml-query-engines}, while others use XML as a means of data exchange
by implementing of middleware to query and extract query data
\cite{sn-xml-middleware}. NetBEAMS is one of such sensor network that contains
an XML-based communication middleware for data exchange among the sensor nodes
\cite{netbeams2009}.

\section{Data Provenance: How Collected Data is Described?}
\label{sec:sn-provenance}

Most of sensor networks implementations use the collected data from the sensor
devices after they perceive the environment. As shown by
\cite{sn-data-model-survey}, different data models contains their pros and
cons to represent data. However, one crucial feature in sensor networks is the
description of it collected data, as \cite{sn-provenance} discussed the
properties and nature of sensor networks generated data. First, the device's
sampling are usually in raw format or without obvious names. In order to
provide search over the collected sensor data, the data needs first to be
described and indexed, and this process can be related to Data Provenance
\cite{db-provenance}. Needless to say, Data Provenance is the
Metadata\footnote{Data about data} about the collected sensor data.

An example of Data Provenance application is the use of aggregation of data
over time to estimate the changing effects of the weather in a given region.
For this reason, the collected data needs descriptions or annotations that can
provide enough information to track changes, that is, the history of how and
when a given data has changed. The following should be the essence of a
provenance-aware collected data according to \cite{sn-provenance}.

\begin{itemize}
  \item \textbf{what was collected?}: it can be merely one or more low-level
  observations,  which is a set of values with or without a description;
  \item \textbf{When was the data collected?}: is the time when the data was
  collected,  as most sensor data must be used depending on its existence;
  \item \textbf{Where was the data collected?}: it is typically the value of 
  the of the location from where the data was collected. However, it is vital to
  note that the use of this metadata depends on the applicability of the data.
\end{itemize}

While Data Provenance provides basic mechanisms to describe sensor networks
data in general, others see the importance of the time space of sensor
networks. \cite{sn-time-series} describes strategies to used to cluster
distributed sensor data as time series of data, as well as to decrease the
size of data streams being transmitted in real-time sensor networks when these
collected data does not need to be transferred \cite{sn-data-reduction}.

Finally, an indispensable application of collected sensor data is the so-called
blackbox of an aircraft, which carries the collection of raw data collected
from a group of sensors placed in different places of the aircraft, as
described by \cite{sn-exemple-blackbox}. The author describes the data is used
in a process of data recovery and analysis in order to identify the probable
reason of an aircraft accident, by using the time, and location of the
aircraft, along with the independent variables from the sensors.

\subsection{What was collected?}

Different sensor devices will produce different attribute sets regarding a
particular observation. In general, these data sets does not have distinct
names, which leads to difficulties when indices are required in order to make
the data searchable. Specially for the end of data archival, as described
earlier in this chapter, the collected data needs to receive and identity,
which identifies the data by itself in a regular database
\cite{relational-model}. Equally, \cite{sn-provenance} described that
traditional sensor networks uses the collected data for analysis long after
the attribute set has been collected, usually before it is categorized and
indexed. Annotations and tags are also names given by \cite{sn-provenance} as
means to describe collected sensor data.

In a nutshell, data is described by the name of the observation, when that is
not provided. For example,

\begin{itemize}
  \item \textbf{temperature} = 73.45: this is the value of the collected
  observation, which in general, the attribute name ``temperature" is not
  provided in the data stream;
  \item \textbf{temperature-scale} = ``Fahrenheit": the metadata regarding the
  temperature attribute, which describes the scale of the temperature;
  \item \textbf{device-version} = 5: the metadata that describes the version of
  the sensor device, since sensors can be replaced, and as a consequence, data
  accuracy may be as well be changed.
\end{itemize}

\subsection{When was the data collected?}

In order to use historical sensor data, the use of a point in time is the most
prominent feature of data, as it is considered a pivotal metadata for different
types of sensor networks such as an aircraft the blackbox
\cite{sn-exemple-blackbox}. Other projects classify data in time-series
\cite{sn-time-series-example}, while others provide approaches to decrease
the creation of data based on clustering data before they are sent to the sink
\cite{sn-time-series}.

In different areas, temporal databases or time-series are the ones that take
the real-world time into consideration. In general, data in a temporal database
can be classified by its dimension, as surveyed by \cite{db-temporal}:

\begin{itemize}
  \item \textbf{Valid time}: it is the point in time during which a fact
  happened with respect to the real world. This value is only used when data is
  captured, and cannot be changed because of its nature, as well as placing a
  restriction over Update operations;
  \item \textbf{Transaction time}: is the point in time at which a fact is
  collected, transferred or processed in the database. This value is used
  during Insertion operations, as well as in Update ones.
\end{itemize}

When both temporal dimensions are applied, the data is considered to be
Bitemporal when \cite{db-temporal}. Usually, the data type of the valid and
transaction time columns are straightforward timestamps, with any supported
precision. For example, the following set of attribute and values shows a
temperature of 70 degrees that was observed on ``November 15, 2009'', but
collected only 3 days later.

\begin{itemize}
  \item \textbf{temperature} = 70;
  \item \textbf{fact-time} = ``Nov 15, 2009 10:45 PM PST''
  \item \textbf{transaction-time} = ``Nov 18, 2009 11:11 PM PST''
\end{itemize}

When it comes to Data Provenance in sensor networks, \cite{sn-provenance}
shows that the collected data for archival purpose is usually saved in 
self-described flat-files by file-names, containing complex naming conventions
that contain the timestamp information about the data. For this reason, time
is a central metadata needed in any data model used, since it helps
identifying data uniquely for a given data set.

\subsection{Where was the data collected from?}

Depending on the application of the collected sensor data, the location in
which the data was collected may be needed. For example, the geographic
location of the collected data is critical to identify the specific
GPS\footnote{Global Positioning System} coordinates of the sensor device
\cite{sn-ex02}. Similarly, the position of any  event captured in a black-box
of an aircraft carries is also used as metadata for the event. It will either
depend on the sensor device provide such sampling attribute, or entered
manually on the database.

A correlation between time and location is presented by \cite{sn-time-series}
as a noteworthy feature to reduce the traffic generated in a sensor network by
clustering similar data when the sensor devices are placed in concentrated 
environment.