### 2009-12-08 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{DSP Data Persistence: Component Design and Architecture}

In order to provide information to users, the RTC scientists use the
\emph{OPeNDAP} \cite{opendap} server to distribute collected data com the
\textit{SF-BEAMS} sensor network over the Internet. As discussed in the previous
chapter, the database system \emph{mongoDB} was selected for the evaluation of a
backend system for NetBEAMS. Unlike other projects reviewed in the literature, mongoDB has
never been used as a selected database for data collected from sensor devices.
In this way, the design of the proposed solution follows the requirements and
specifications detailed in chapter 4. As described before, NetBEAMS provides a
component-based infrastructure that automates the operation of SF-BEAMS sensor
network. The proposed solution presented in this chapter is based on the
NetBEAM's DSP Platform architecture, as described in chapter
\ref{chap:netbeams-overview}. The following sections describe the scenarios for
persistence, analyze the requirements captured from the RTC staff, and describe
the design and architecture of the proposed solution.

Technically speaking, this chapter includes detailed UML diagrams regarding the
different point of views of the system, following the conventions defined at
\cite{uml}.

\section{Business Process Analysis}
\label{sec:business-process-analysis}

As described in section \ref{chap:netbeams-overview}, the DSP Platform was
designed to support SF-BEAMS, an environmental sensor network based in San
Francisco Bay, composed by different types of sensors. As shown in in Figure
\ref{fig:dsp-persistence-business-process}, the event of data collection is
started by the DSP Component deployed on each Gumstix system based on the
sampling rate frequency value. When the DSP Client component finishes the data
extraction process, a new process starts at the client side. First, each of the
co-located DSP Client processses the collected measurements and transforms
them into the alleged DSP Message, as shown in section \ref{sec:dsp-message}, 
emboding the actual collected data from the sensor devices. In this way, an 
instance of a DSP Measurement Message is added into the local temporary
outbound queue that waits to be activated. Then, as an asynchronous data
transmission event occurs within a specified rate, transferring all the queued
messages to be processed by a DSP Server host. This server host is centralized
and recives all the DSP Messages, to be forward to different messages.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Business-Diagram}
  \caption{UML Business Process Model diagram - Adding Persistence for NetBEAMS}
  \label{fig:dsp-persistence-business-process}
\end{figure}

Once the DSP Server processes the received DSP Message, it uses the server
configuration to transmit the collected data to the Measurements
Database. As a matter of fact, the collected data from sensor devices are basic
properties with keys and associated numerical values, described in section
\ref{sec:sfbeams}. According to the descriptions of the DSP Platform and the
DSP Messages structure, the collected data from the sensors is parsed and
mapped to a POJO\footnote{Plain Old Java Object} or it can be serialized in
XML \cite{xml}. Similarly, the collected data contains attributes that
uniquely identifies them, along with provenance metadata as spatio-temporal
attributes.

\section{Requirements Specification}

After analyzing the main data collection scenario and reviewing the
requirements defined by the RTC staff, this section enumerates the main
functions that the proposed solution must perform.

\subsection{Functional Requirements}
\label{sec:use-cases}

The primary goal of system to be implemented is related to the collected data
reuse. That is, allow users to have easy access to the collected data from
\emph{NetBEAMS}. The main motivations of the proposed system is that it must
give different types of users access to the collected data without requiring
the use of SQL. Rather, it must provide a programming language abstraction
that might make the access to those data easier \cite{sn-programming-language}.
Furthermore, it must use the selected technology, mongoDB, using the
properties defined by the system such as the use of a schema-less database
systen. In this way, the designed functions for the solution, as depicted in
the UML Use Cases Diagram of
Figure \ref{fig:DSP-Data-Persistence-UseCases-Diagram-Users}, are summarized as
follows:

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-Users}
  \caption{UML Use Case diagram for Persistence Functions}
  \label{fig:DSP-Data-Persistence-UseCases-Diagram-Users}
\end{figure}

\begin{itemize}
  \item \textbf{Search Measurements by Sensor Properties}: users of the
  collected data must be able to search data through the properties of the
  sensors. This is related to identity principle of what was collected
  as provenance \cite{db-provenance};
  \item \textbf{Search Measurements by Sensor Location}: users of the collected
  data must be able to search data by a given sensor location, as referred to
  the provenance from where the data was collected;
  \item \textbf{Search Measurements by Data collection date}: users of the
  collected data must be able to search data by a given date, or ranges of
  start and end date;
  \item \textbf{Annotate Existing Measurements}: users of the collected data
  must be able to annotate any existing measurements, that is, they must be
  able to add more metadata such ``tags'';
  \item \textbf{Export Data to Spreadsheets}: users of the collected data must
  be able to export the collected data to spreadsheets.
  \item \textbf{Export Data to OPEnDAP format}: users of the collected data
  must be able to export the collected data to the OPEnDAP format;
  \item \textbf{Access the data through API, Programming Languages}: users of the
  collected data must be able to search data by the use of different
  programming and scripting languages of their preference;
  \item \textbf{Remove Undesired Measurements}: users of
  the collected data must be able remove any collected data.
\end{itemize}

On the other hand, the functions performed by the solution must be designed
conformed the specifications of the Data Sensor Platform (DSP) architecture,
described in section \ref{sec:netbeams-architecture}. In this way, the
solution must be a Data Producer, implementing the following functionalities,
as shown in Figure \ref{fig:DSP-Data-Persistence-UseCases-Diagram-System}:

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-System}
  \caption{UML Use Case diagram for the Persistence Component}
  \label{fig:DSP-Data-Persistence-UseCases-Diagram-System}
\end{figure}

\begin{itemize}
  \item \textbf{Update Component Properties}: as described by the DSP Component
  architecture, the new DSP component must be able to have its properties
  updated. These properties must include the rate of how long it takes to the
  component flush messages;
  \item \textbf{Update Database Properties}: the component must be able to
  update the database settings such as IP Address and Port of the service;
  \item \textbf{Collect DSP Measurement Message}: as the component is designed
  to be a Data Consumer, it must be able to receive and temporarily retain the
  collected messages from the sensors;
  \item \textbf{Categorize Measurement Message}: as the DSP Message is
  delivered to the new component, it must categorize the message by its
  properties such as transaction time, originating DSP Message ID, etc;
  \item \textbf{Save Messages into the Database}: the main goal of the new
  component is to send the payload of the DSP Messages to the database in use,
  in this case, mongoDB.
\end{itemize}

\subsection{Non-Functional Requirements}

Different constraints were required to execute the functional requirements
described in the previous section, that is, non-functional requirements that
the system must follow:

\begin{itemize}
  \item The solution must maintain the received DSP Measurement Messages
  in-memory for a specific period of time in order to decrease the write load
  on the database (flusher rate);
  \item The Persistence storage system must be able to cope with the Data
  Volume produced by SF-BEAMS in a daily bases;
  \item The Persistence storage system must be able to scale without too much
  architectural changes, or schema changes;
  \item The Persistence storage system must be able to provide access to the
  persisted data using an abstraction similar to programming languages, as
  the authors of \cite{sn-programming-language} proposed;
  \item The Persistence storage must be able to 
\end{itemize}

\section{Data Model Design}
\label{sec:dsp-persistence-data-model}

MongoDB is defined as a Document-Oriented Database system, that uses the
abstraction of the Key-Value-Pair data model. In this way, mongoDB does not
require the creation of a schema. However, the definition of how to describe
the collected data must be provided, as they define the keys to the system.
Similarly, in order to describe the keys, the specifications details the use of
Data Provenance as guidelines to the description of the collected data.
Therefore, the key-value pairs of any instance must include the following
properties:

\begin{itemize}
  \item \textbf{Locality}: properties that describe the location of the data.
  For example, the host machine name or IP address, the GPS coordinate
  \cite{gps};
  \item \textbf{Identity}: what uniquely identities the data among the others
  produced by other sensor devices in the network, as for instance a
  UUID\footnote{Universally Unique Identifier};
  \item \textbf{Time-Dimensions}: when the data was ``sensed''
  \cite{db-provenance} and when the collection transaction happened
  \cite{sn-time-series}.
\end{itemize}

After analyzing the DSP Platform and the structure of the DSP Messages in
chapter \ref{sec:netbeams-architecture}, the following set of keys were designed
as be part of the data model, defined as the ``key segment'':

\begin{itemize}
  \item \textbf{message\underline{ }id}: Defines the DSP Message IP that
  collected the measurement message. This can be used to track which component generated, etc;
  \item \textbf{sensor}: Defines the properties of the sensor device;
    \begin{itemize}[label=\textbullet]
        \item \textbf{ip\underline{ }address} as the IP Address of the sensor
        device that produced the data;
        \item \textbf{location}: defines the location of the sensor device;
            \begin{itemize}[label=\textbullet]
                \item \textbf{latitude} the GPS latitude coordinate of the
                position of the sensor device, captured from references of the device;
                \item \textbf{longitude}: the GPS longitude coordinate of the
                position of the sensor device.
            \end{itemize}
    \end{itemize}
  \item \textbf{time}: identifies the time dimensions about the collected data;
    \begin{itemize}[label=\textbullet]
      \item \textbf{valid}: it is extracted from the container's date and time
      of creation, which was collected during the sensor device's reading;
      \item \textbf{transaction}: it is extracted from the DSP message
         container creation time and it is used to identify when the
         transaction occurred;
    \end{itemize} 
\end{itemize}

While the ``Key Segment'' is the set of keys formed by the Data Provenance
data, the ``Value Segment'' stores the observations from the sensor
devices. In this sensor, the observation key is defined as follows::

\begin{itemize}
  \item \textbf{observation}: this key stores the values of the document, and
  will have every different property of a given sensor device. For instance, the
  data key for the properties of the YSI device will contain the set of properties 
  defined for the device, such as battery, temperature, salinity, among other,
  as shown in Listing \ref{file:mongodb-ysi-data-format}.
\end{itemize}

\section{High-Level System Architecture}

Considering the business analysis described in section
\ref{sec:business-process-analysis}, the addition of a data persistence layer
to the DSP Platform can be achieved by adding Data Consumer (DC) as described
in section \ref{sec:netbeams-architecture}. Based on the Requirements
specification described earlier, this component must be responsible for
providing the processing of the DSP Measure Messages, extracting the payload of
the body of the message and converting it to the data model defined in the
previous section. Finally, the data can be transmitted to the chosen database
system. Figure \ref{fig:NetBEAMS-Persistence-Server-Node-Components} provides
an architectural view of the DSP Server node with the loosely-coupled
components, highlighting the inclusion of a new DSP Component called ``DSP Data
Persistence''.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/NetBEAMS-Persistence-Server-Node-Components}
  \caption{UML Components diagram of the DSP Data Collector}
  \label{fig:NetBEAMS-Persistence-Server-Node-Components}
\end{figure}

As it is depicted in figure
\ref{fig:NetBEAMS-Persistence-Server-Node-Components}, the DSP Data
Persistence component can be added into the system as an independent
plug-and-play component. In addition, the database system and the necessary
drivers are also added without any changes to the existing DSP infrastructure.
In this way, a matching rule must be provided. First, as DSP Messages are
delivered to the server through the DSP Wire Transport Server, as shown by
flow 1, the DSP Platform is responsible to forward a copy of the Measurement
messages directly to the DSP Data Persistence as shown by flow 2. Then, the
sole responsibility of the DSP Data Persistence is to transfer the received
measurement messages to the database system by using the connection drivers,
as depicted by flow 3.

The following sections focus on the DSP Data Persistence component design in
the context of an OSGi bundle \cite{osgi}.

\section{DSP Data Persistence: OSGi-DSP Bundle Design}

The new proposed Data Persistence component must follow the DSP requirements
for a new DSP Component, which specifies the design-pattern of Data
Producer/Consumer. In a nutshell, the design of a DSP Component maintains the
dependency among 3 different major Java packages: it depends on 2 existing
packages, namely the OSGi Framework and the DSP Platform, as well as an
additional package for the database access driver API, as shown in the UML
Package Dependency Diagram of Figure
\ref{fig:DSP-Data-Persistence-Packages-Dependency}.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Packages-Dependency}
  \caption{UML Components diagram of the DSP Data Collector}
  \label{fig:DSP-Data-Persistence-Packages-Dependency}
\end{figure}

The following sections details each step of the DSP Data Persistence activation
and parallel execution of the contained classes.

\subsection{DSP Data Persistence Activation and Message Delivery}

The DSP Data Persistence component starts with an OSGi Activator class, and the
implementation of the DSP Component, resposible to provide services to the
platform. Based on these specifications, the following set of classes in
figure \ref{fig:DSP-DataPersistence-Activator-Class-Diagram} shows the UML
Class Diagram implemented for the component.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Activator-Class-Diagram}
  \caption{UML Class diagram for the DSP Data Persistence Activator and Component}
  \label{fig:DSP-DataPersistence-Activator-Class-Diagram}
\end{figure}

The DSP Data Persistence Activator is the main class responsible for the
activation of the DSP Data Persistence component, as it extends the OSGi class
BundleActivator and maintains a reference of the DSP Data Persistence class.
Furthermore, the activator maintains references to the classes BundleContext
and ServiceReference both from the OSGi Framework. The former is responsible
for maintaining the link with the OSGi Framework, while the latter is used to
register the DSP Component as a service. Therefore, it is clear that the class
DSP Data Persistence Activator is the main communication interface between the
DSP Platform and the OSGi Framework.

While the DSP Data Persistence Activator is responsible for the orchestration
of the classes during the execution on the OSGi Framework, the class DSP Data
Persistence extends from the DSP Platform class DSPComponent. For this reason,
this class inherits all the behavior described on \ref{sec:netbeams-architecture}
and can be in the roles of Data Consumer and a Data Producer.

\begin{itemize}
  \item \textbf{DSP Data Persistence as Data Consumer}: receives any measurement
  message from remote sensors;
  \item \textbf{DSP Data Persistence as Data Producer}: transforms any
  measurement message received into a format that must be ready to be saved on the
  database, based on the description of the data model in section
  \ref{sec:dsp-persistence-data-model}.
\end{itemize}

The behavior of the activator class follows the specifications of the
OSGi Framework and, for this reason, instances of the DSP Data Persistence
Activator can be in the states described on the UML State Diagram depicted in
Figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}. It is important to
note that the activator is an extension of the OSGI class BundleActivator, as
well as being responsible for stopping or starting the the DSP Data Persistence
component instance. In this way, following the specifications of section
\ref{sec:netbeams-architecture}, the DSP Data Persistence Platform can be
initialized as described by the UML Sequence Diagram of Figure
\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}:

\begin{enumerate}
  \item During the DSP Platform activation, it will get the name of the
  persistence component from the configuration artifact config.xml as seen in
  listing \ref{file:dsp-config.xml};
  \item After being selected based on the configuration priority, the DSP
  Bundle artifact is identified and installed by creating an instance of the
  class DSP DataPersistence Activator and making a call to the method start();
  \item During the activation, an instance of the DSP Data Persistence
  component is created and registered as an OSGi Service;
  \item Upon registering the DSP Data Persistence component, the DSP Platform
  "listens" the event serviceUpdate() and, and triggers the last operations to
  bootstrap the DSP Component:
   \begin{enumerate}
      \item Initializes the component by calling the method initComponent();
      \item Starts the component by calling the method startComponent();
      \item Bootstrap messages are delivered to the component.
   \end{enumerate}
\end{enumerate}

Any configuration parameter required by a DSP Component is sent during its
bootstrap process through the use of an optional DSP Update Message, as shown
in Listing \ref{file:dsp-data-persistence-bootstrap.xml}. The initialization of
the DSP Component Activator and the DSP Component registration are also shown
in Figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
  \caption{UML Sequence diagram describing the DSP Data Persistence bundle activation}
  \label{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
\end{figure}

As soon as the gate "Start DSP Activator Gate" reaches the DSP Data
Persistence instance, it first creates the unique instance of the class DSP
Data Persisence and uses that instance to register it as the OSGi service
through the call to the DSP Platform helper class ActivatorHelper. At this
point, the DSP Data Persistence has been installed and is in the state
"Active".

\subsection{Delivering Bootstrap and Measurement Messages}

Although the DSP Data Persistence component has been instantiated as of Figure
\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}, the
process of delivering messages to the DSP Data Persistence has yet to be
completed. As defined by the requirements, the DSP Component must execute its
function in a concurrent fashion, using configuration parameters during its
bootstrap process. The class DSP Data Flusher is responsible for running as a
"worker thread", whose function is to verify if there are messages to be
flushed into the Database with the class Transient Persistence Layer. Figure
\ref{fig:DSP-DataPersistence-Flusher-Classes} describes the UML Class Diagram
for the main classes of this functionality.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Flusher-Classes}
  \caption{UML Class diagram with the DSP Data Flusher worker thread}
  \label{fig:DSP-DataPersistence-Flusher-Classes}
\end{figure}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Mongo-Classes}
  \caption{UML Class diagram with Persistence Service for MongoDB}
  \label{fig:DSP-Data-Persistence-Mongo-Classes}
\end{figure}

The DSP Data Flusher is a thread that is designed to be in two different
states as depicted in Figure
\ref{fig:DSP-DataPersistence-Flusher-State-Diagram}: RUNNABLE and
TIMED\underline{ }WAITING. In the former state, the DSP Data Flusher will be
contacting the Transient Persistence Layer in order to verify if there are any
message in the outbound queue ready to be transmitted to the Database
service, while the DSP Data flusher will be waiting for the next cycle defined
by the bootstrap parameter TRANSIENT\underline{ }DATA\underline{
}FLUSHER\underline{ }DELAY in the latter state.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Flusher-State-Diagram}
  \caption{UML State diagram for the DSP Data Flusher Worker Thread}
  \label{fig:DSP-DataPersistence-Flusher-State-Diagram}
\end{figure}

Upon receiving a new DSP measurement message, the DSP Data Persistence
component must send its reference to the temporary memory space, as described
by the non-functional requirement in the beginning of this chapter. In this
way, the DSP Data Persistence component maintains a dependency on the Singleton
class TransientPersistenceLayer, which is responsible for keeping track of the
messages that were received by the running component. Finally, in order to add
the DSP message into its aggregated set, it first transforms the message into an
instance of the class PersistentMessageUnit. This class carries references to
the class \emph{SensorLocation}, responsible for handling the location of the
sensor device, and consequently, the physical origin of the message.

A DSP Message is deliveried to the DSP Broker with the intent of being
forwarded to the data consumers, which are configured to receive that
specific message. This process is described by Figure
\ref{fig:From-DSPWireTransport-Server-To-DSP-Broker}. Similarly, continues the
steps of the DSP Broker from the gate "From DSP WireTransport Server to Broker
Gate", showing the steps taken to transfer a DSP Message to the Singleton
class TransientPersistenceLayer, responsible for maintaining a queue of
messages waiting to be sent to the database system.

According to the DSP Broker model described in section \ref{sec:dsp-message}, 
the DSP Matcher needs to be updated with the addition of a matching rule that
filters a copy of any Measurement Message to the DSP Data Persistence component
designed. As it is shown in figure
\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}, the worker
Thread DSP Data Flusher is the link between the Transient and Persistent
layers, as it depends on the references from the classes
TransientPersistenceLayer and the DSPMongoCRUDService. The former maintains
DSP Messages in memory wrapped by an instance of PersistenceMessageUnit and
the latter is responsible for sending the MessageContent from the DSPMessage
body to the Database.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSP-Broker-To-DSPDataPersistence-General-Sequence}
  \caption{UML Sequence Diagram - Flow 1: From the DSP Broker to the Data
  Persistence Component}
  \label{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}
\end{figure}

The DSP Broker retrieves the list of matching rules by contacting the Matcher.
As the DSP Platform have already loaded all the Matching rules, a copy of a
given DSP message is delivered to the DSP Data Persistence component. However,
the sequence diagram shows two different stages of reception:

\begin{itemize}
  \item A bootstrap DSP Update Message is delivered to the DSP Data Persistence
  component when the DSP Platform is initializing the component. In this way,
  the DSP Data Persistence can initialize the DspDataFlusher worker thread to
  flush temporary messages into the database at the rate of
  TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY in seconds,
  among others. Listing \ref{file:dsp-data-persistence-bootstrap.xml} shows an
  implementation of a bootstrap message for the DSP Data Persistence;
  \item Any DSP Measure Message can be persisted by the DSP Data Persistence
  component. The component will add the message to a temporary memory location
  to decrease I/O on the database, when the DSP Broker delivers a DSP
  MeasureMessage to this component, as depicted by the gate "create
  PersistentMessageUnit Gate". This last gate describes the simple procedure
  to add the message into the transient persistent layer. Listing
  \ref{file:dsp-message-serialized-ysi}.
\end{itemize}

The class PersistenceMessageUnit is the major transient persistence unit that
carries references regarding the originating DSP Message that was collected by
the DSP Data Persistence component. It is composed by a SensorLocation and
contains a reference to the PersistentMessageState enumaration. The former
identifies which sensor produced the collected data from the DSP Message by the
IP address, while the latter identifies whether the PersistenceMessageUnit has
been saved into the Database or not as depicted by the UML State Diagram in
Figure \ref{fig:PersistentMessageState-Diagram}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/PersistentMessageState-Diagram}
  \caption{UML State Diagram for the instance of the class
  PersistentMessageState}
  \label{fig:PersistentMessageState-Diagram}
\end{figure}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence}
  \caption{UML Sequence diagram - Adding a DSP Message into the Transient Persistence layer}
  \label{fig:From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence}
\end{figure}

According to the data model specifications defined in section
\ref{sec:dsp-persistence-data-model}, the persistence model must carry
information regarding which sensor device produced the collected data. For
this reason, an instance of the class SensorLocation carries the information
about the IP address and and optional GPS location from the component that
originally produced the DSP Message. Then, an instance of the class
PersistenceMessageUnit is created at the state ``TRANSIENT'', as detailed
on UML State Diagram in Figure \ref{fig:PersistentMessageState-Diagram}.

\subsection{Flushing data into the Database}

The last stage sequence described in the UML Sequence diagram of figure
\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} depicts the
DspDataFlusher worker thread continuously waking up at the rate of
TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY. Its purpose
is as simple as to iterate over the list of PersitenceMessageUnit on the state
of TRANSIENT, and send each of them to be saved on the persistence storage.
Figure \ref{fig:DSP-Data-Persistence-Classes} shows the participating classes
of the component.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Classes}
  \caption{UML Class diagram for the DSP Data Persistence component.}
  \label{fig:DSP-Data-Persistence-Classes}
\end{figure}

Finally, this sequence of events is continued by the "insert
PersistenceMessageUnit Gate", which triggers the persistence service from the
chosen database system as shown in figure
\ref{fig:From-Insert-PersistentMessageUnit-to-mongoDB}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Insert-PersistentMessageUnit-to-mongoDB}
  \caption{UML Sequence Diagram - MessageContent instance saved by the class
  mongoDB service}
  \label{fig:From-Insert-PersistentMessageUnit-to-mongoDB}
\end{figure}

In order to build the document key and value segments, as described on section
\ref{sec:dsp-persistence-data-model}, the reference to the DSP message is
used, as the value key is built by extracting the DSP Message Content from the
body of the DSP Message. Similarly, the key is built by gathering information
from the DSP Message and the PersistenceMessageUnit class.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Creating-Value-Segment-Sequence}
  \caption{UML Sequence diagram showing the creation of the value document segment}
  \label{fig:From-Creating-Value-Segment-Sequence}
\end{figure}

Finally, the document key segment is prepared as follows in figure
\ref{fig:From-Creating-Key-Segment-Sequence}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Creating-Key-Segment-Sequence}
  \caption{UML Sequence diagram showing the creation of the key document segment}
  \label{fig:From-Creating-Key-Segment-Sequence}
\end{figure}

\section{DSP Data Persistence Deployment}

The deployment of the persistence layer is comprised by the addition of two
new components to the DSP Platform: the DSP Data Persistence and the MongoDB
Database Server. The former is responsible for capturing any data
produced by sensors in the data sink, as described in the previous sections of
this chapter. On the other hand, the latter is where the data is going to be
stored, as a result of the technology selection on the previous chapter.

The deployment of the DSP Data Persistence follows the specifications of how to
deploy a DSP Component on an OSGi Framework. For the development of this work,
the Knopflerfish OSGi Container was used, as described in section
\ref{sec:dsp-data-persistence-implementation}. In contrast, the
deployment of the mongoDB can be accomplished in different approaches, as
summarized:

\begin{itemize}
  \item \textbf{Single Server}: a regular deployment of a database system as it
  is done in general applications, where data is managed by one single process;
  \item \textbf{Sharded Server}: a deployment approach where data is sharded
  into different servers.
\end{itemize}

The single server approach maintains all data located in a simple space.
Distributed Systems applications such as Master-Slave, Data Replication, etc
can be used in this approach for scalability
\ref{fig:DSP-Data-Persistence-Deployment-Single}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.7]{../diagrams/DSP-Data-Persistence-Deployment-Single}
  \caption{UML Deployment diagram with a single server}
  \label{fig:DSP-Data-Persistence-Deployment-Single}
\end{figure}

However, given that mongoDB provides support to Shards, then the implementation
of a Data-Centric approach can be done by deploying different mongoDB instances
in a distributed cluster, where each node can hold a given type of data. The
specification of what types of data should be stored in each of the cluster
nodes is defined by a shard-key.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.7]{../diagrams/DSP-Data-Persistence-Deployment-Sharded}
  \caption{UML Deployment Diagram for a sharded server}
  \label{fig:DSP-Data-Persistence-Deployment-Sharded}
\end{figure}

All in all, the design of the DSP Data Persistence uses the DSP Component
specification defined by the DSP Platform. First, the analysis of the business
process helped understing the requirements of the system. Similarly, the data
model design followed the specifications defined in chapter 5. In this way,
each of the layers of the system were described using UML Diagrams, having
partial details regarding the OSGi implementation. Next chapter entails the
experiments conducted to evaluate the designs discussed in this
chapter, walk-through the experiments conducted to verify the implementation
using mongoDB.