### 2009-12-08 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Experimental Results: Correct Behavior and Performance}

As discussed in the previous chapter, the implementation of a data persistence
layer for NetBEAMS followed the requirements defined in chapter 4 having in
mind the classification of the SF-BEAMS sensor network and the software
infrastructure provided by the Data Sensor Platform. As a result, the
implementation was developed using the infrastructure provided by the online
version-control system repository from NetBEAMS, whose documentation is
detailed in section \ref{sec:dsp-data-persistence-implementation}.

First, this chapter details the experiment setup designed to evaluate the
proposed implementation, showing the measurement results collected from the
simulation log and based on the user experience while reusing the collected
data. Last, but not least, the pertinent discussion of this work is devided up
into different sections, each of them analyzing different aspects related to
data persistence for NetBEAMS, amongo others.

\section{Experiment Setup}

The experiment was designed to support different scenarios of a regular use of
the mongoDB system infrastructure, simulating the activities of data collection
from the SF-BEAMS network using the NetBEAMS infrastructure. First, the setup
for the data collection of randomly generated data was implemented, which
directly uses NetBEAMS already existing classes. Then, the implementation of
the CRUD scenarios were implemented in different programming languages, from
the data insertion scenario using the Java Programming Language, and then the
remaining operations using the Javascript Scripting language \cite{javascript}
and the Python Programming Language \cite{python}. Similarly, an open-source web
application, developed in the Ruby Programming language \cite{ruby}, was also
used. The experiment scripts were writen using Shell Bash Script Language
\cite{bashshell} and is responsible for orchestrasting the execution of one
single experiment ``run''.

\subsection{Experiments Artifacts}

Different artifacts were created to perform the different experiments
operations. The Create function was developed in Java, without the interversion
of the DSP or OSGi, isolating only the functionalities of provided by CRUD
service class DSPMongoCRUDService (Listing \ref{file:dsp-mongo-service}),
responsible for the connection handling with mongoDB, as well as the insertion
operation from NetBEAMS. As described in section \ref{sec:dsp-details}, the
current version of NetBEAMS supports the data collection for the YSI Sonde
device, among others used for demo purposes such as a software sensor that
captures the mouse activities over a window. The documentation of both data
types is seen in section \ref{sec:dsp-payload-implementation}. In this way, the
DSP Data Persistence, as described in the previous section, supports both data
types to be persisted.

In order to exercise mongoDB with the same data collected from NetBEAMS, the
existing Java class TestSondeData (Listing
\ref{file:random-ysi-data-generator}), was refactored to provide a random
instance generator for the YSI Data handler implemented in the Java class
SondeDataType (Listing \ref{file:main-ysi-data-handler}). When the Insert
function is executed, the created instances are transferred from the JVM to
the mongoDB, and stored in the collection ``SondeDataContainer''. Specific
details about the execution of mongoDB are depicted in section
\ref{sec:mongodb-deployment}. Finally, the implementation of the experiment
in the Java class DSPMessageToMongoDBExperiment (Listing
\ref{file:experiment-dsp-java-executor}) was responsible for managing the
creation and insertion of any workload composed by instances of the class
SondeDataType in a bulk manner.

Similarly, the support for the Retrieve, Update and Delete operations against
the inserted workload were implemented using Javascript, as shown in Listing
\ref{file:experiment-scenarios}, respectively by the function calls ``find'',
``update'' and ``remove'' over the collection ``SondeDataContainer''.
Similarly, in order to implement a customizable export system that converts
from mongoDB into the format OpenDAP, a simpler prototype script was developed
in Python, as shown in Listing \ref{file:experiment-export-python}. As a result
of the execution of this script, the exported artifact has the same format as
described in Listing \ref{file:rtc-ysi-opendap}.

On the other hand, the execution of the main experiment shell script requires
the use to provide the size of the workload, as shown in listing
\ref{cmd:run-persistence-experiment}.

\lstset{label=cmd:run-persistence-experiment,caption=Running main experiment
shell script}
\begin{lstlisting}
marcello@netbeams-dev:~/workspaces/netbeams/versions/v2/persistence$ ./run-persistence-experiment 
#######  NetBEAMS Experiments - Persistence on MongoDB  ########

Usage: ./run-persistence-experiment X, where X is the size of the workload to be inserted from NetBEAMS YSI Data Handler to mongoDB
\end{lstlisting}

\subsection{Used Workload}
\label{sec:workload}

The workload selected for the experiments reflects the current use of the data
at the SF-BEAMS infrastructure. In this way, the experiment runs for 5
different times performed to investigate different scenarios to be described.
In this way, this number is related to the total number of sensor devices
reported to be in operation, as discussed in section \ref{sec:sfbeams}. In
summary, the volume of the workload prepared after the experiment setup
is as follows:

\begin{itemize}
  \item \textbf{First Round}: 483,840 documents, worth the production of the 
  YSI Sonde device during one year, at the rate of 1 observation per minute;
  \item \textbf{Next Rounds}: 483,840 * 4, or 2,419,200 documents, representing
  5 different YSI devices similar to the first round.
\end{itemize}

\subsection{System Environment}

Two different system environments were prepared: one for the External Storage
environment, or Single Server; and another for the Data-Centric Storage, or
Distributed Server. Since the single server is a regular centralized database
system, mongoDB just requires the execution of the database process ``mongod'',
which is started by specifying the directory where the inserted data will be
managed. On the other hand, nonetheless, the distributed server setup requires
a different list of processes (see section \ref{sec:mongodb-user-experience}):
one ``mongod'' process for each mongoDB shard running on the different mongoDB
cluster, one ``mongod'' process to manage the cluster metadata, and one
``mongos'' process, which is the main one responsible to orchestrate the
cluster. In conclusion, the single system environment is synmmetric to
the External Storage used to persist the collected data, while the clustered one
is simulates the Data-Centric Storage one.

The simulation of both the single and distributed system used both a Mac OS 10
Snow Leopard 64bit, with a total of 8GB of physical memory. However, in order
to simulate different shards in different machines, the use of a VirtualBox
\cite{virtualization} provided the use of guest operating systems to run and
act as the mongoDB shards. In this way, two different versions of the Ubuntu
Linux were used: one 8.04 32bits and another 9.04 64bits. The configuration
used for the single server only uses the process ``mongod'' to be started,
while the clustered setup must be enabled in the database, as shown in
Listing \ref{file:experiment-setup-mongoshards}, Listings
\ref{file:setup-os-shards-step1} and \ref{file:setup-os-shards-step2}.

\subsection{Planned Scenarios}
\label{sec:exp-scenarios}

In order to verify the feasibility of the solution proposed, all CRUD
operations were exercised through the implementation of the practical
scenarios based on the use cases described in section \ref{sec:use-cases}. In
brief, the use cases were chosen to run with random values such as follows:

\begin{itemize}
  \item (R0) Insert a volume of collected data from sensors using NetBEAMS
  service compatible with 1 year of observations;
  \item (R1) Find all documents whose observations were collected between two
  different dates, say the days between December 8 and 15 of 2009;
  \item (R2) Find all documents whose observations were collected from a given
  known sensor device such as one whose IP address is equals to "192.168.0.102";
  \item (R3) Find all documents whose observations contained specific values
  read from the environment, say Salinity equals to 0.01 and the water temperature
  equals to 46.47;
  \item (U1) Update all the documents whose observations were collected during
  the day of December 2nd, 2009, by adding a tag = ``oil spill'';
  \item (D1) Remove all the documents whose observations were collected on
  December 7th, 2009.
  \item (R4) Export the produced data of the week to CSV files, using the same
  tabular format sequence used in the files distributed by SF-BEAMS using the 
  the OPEnDAP format, as in Listing \ref{file:rtc-ysi-opendap}.
\end{itemize}

\section{Measurement Results}
\label{sec:exp-measurements}

As described in the previous section, the design of the experiments were based
on the specifications of a data persistence for NetBEAMS, using the mongoDB
database system. In this way, after setting up the experiment environment,
the execution of the workload was performed against the mongoDB database using
its mechanisms used to manipulate the data generated. It uses the programming
language abstraction to provide access and modifiy the data collected from the
NetBEAMS component. For this reason, this section details these mechanisms uses
to implement each of the scenarios defined in the previous section, associating
numerical response time collected from the logs.

\subsection{Insert Operation}

The first set of measurements collected were the ones regarding the use case
R1, shown in Listing \ref{file:dsp-mongo-service}, defined as the insertion of
instances of the collected data from the sensor devices using Java. The
document follows the specification of the keys designed in section
\ref{sec:dsp-persistence-data-model}. The use of the mongoDB Java driver, which
provides the methods to access the database system and the collection. In this
way, the construction of the keys and respective values were done using a
Hash-like object ``BasicDBObject'' such as the ones shown in the method
``buildKeySegment''. Similarly, the creation of database indexes is performed
using the same approach, as shown in the Java method ``setupMetadataIndexes''.
Finally, the insertion of the documents is performed on the service method
call ``insertPersistentUnitMessageContents'', which delimits the bulk insert
using the transaction management mechanism over the collection. In this way,
different insert times were collected from the logs of the different
experiments as summarized on Table \ref{tab:experiment-insert-avarage}.

\begin{table}
    \begin{center}
        \begin{tabular}{|p{100pt}|p{100pt}|p{100pt}|}\hline
           x & \textbf{Linux VirtualBox Server} & \textbf{Mac OS Host Server}\\\hline 
           \textbf{Single Server} & ~25K docs/sec & ~51K/sec \\\hline
           \textbf{Clustered Server} & ~17K docs/sec & ~39K/sec \\\hline
        \end{tabular}
        \caption{Insertion averages on Virtual and Host in Single or Clustered
        mongoDB Server}
    \end{center}
    \label{tab:experiment-insert-avarage}
\end{table}

After executing the experiments runs, the claimed disk space for
the representation of the randomly generated data is separated into two
different sizes: one related to the amount of data produced by one YSI device,
and another produced by five different YSI devices, as summarized on Table
\ref{tab:experiment-insert-diskspace}.

\begin{table}
    \begin{center}
        \begin{tabular}{|p{100pt}|p{100pt}|p{100pt}|}\hline
        x & \textbf{1 YSI} & \textbf{5 YSIs}\\\hline
        \textbf{Indexed Keys} & 278.33 MB & 1.35 GB \\\hline
        \end{tabular}
        \caption{Amount of disk storage used}
    \end{center}
    \label{tab:experiment-insert-diskspace}
\end{table}

\subsection{Retrieve, Update and Delete Operations}

When it comes to the simulation of the use case scenarios, the implementation
used the mongoDB shell client, which uses Javascript, to implement most of the
operations for ``retrieve'', ``update'' and ``delete'' shown in Listing
\ref{file:experiment-scenarios}. On the other hand, the implementation of the
use case to export the collected data to OPeDAP format used Python, depicted on
Listing \ref{file:experiment-export-python}. The access pattern to use the
collected data is ``db.collection\underline{ }name.operation", where
``collection\underline{ }name'' is the name of the collection used for the
collected data, and in this case, the name of the DSP component that represents
the collected data. Therefore, mongoDB uses dynamic object binding for the
collection object, and executes the CRUD operations as method call with a list
of the needed parameters. Table \ref{tab:experiment-scenarios-mongo-methods}
relates method calls used to implement each of the use cases defined, as
implemented in Listing \ref{file:experiment-scenarios}.

\begin{table}
    \begin{center}
        \begin{tabular}{|c|c|c|c|}\hline
        \textbf{Scenario Type} & \textbf{Language} & \textbf{Method Call} & \textbf{Use Case}\\\hline 
        \textbf{Insert} & Java & db.SondeDataContainer.insert() & R0\\\hline
        \textbf{Retrive} & Javascript & db.SondeDataContainer.find() & R1, R2, R3\\\hline 
        \textbf{Update} & Javascript & db.SondeDataContainer.update() & U1\\\hline 
        \textbf{Delete} & Javascript & db.SondeDataContainer.remove() & D1\\\hline
        \textbf{Export} & Python & db.SondeDataContainer.find() & R4\\\hline
        \end{tabular}
        \caption{mongoDB method calls used for the use cases}
    \end{center}
    \label{tab:experiment-scenarios-mongo-methods}
\end{table}

The execution time for each of the use cases were collected using the logs
collected during the execution of the experiment script manager. The retrieval
operations ran in avarage for ~300ms, while the updates took more than 10
seconds when it involved larger data sets with more than 300 thousand
documents. Similarly, the delete operation took less

\subsection{User Experience}
\label{sec:experiments-user-experimence}

In general, the use experience of mongoDB is related to the use of programming
languages, since most of the methods to create, access and modify the database
is performed using the programming language abstraction of objects and method
calls. The creation of data was the first contact with mongoDB in terms
of data creation in Java. Later, Different methods of data access were used
after the randomly generated data was inserted into mongoDB: using the shell
commands mongo and ``mongoexport'', command-line execution developed in
Python, through the REST Web Services API, and using a web application,
written in the Ruby Programming Language \cite{ruby}, providing the view over
the collected data using a web browser. More details about the execution of
these commands during the experiments is shown in section
\ref{sec:mongodb-user-experience}.

First and foremost, the primary method to access to the database is using the
shell command. Once the user requests an access to the database, he or she can
issue an operation over the collection of documents using the regular the
abstraction programming language, as the shown in Listing \ref{cmd:mongo}. The
CRUD operations are available to the user and can be used at any time to any
database collection. Aggregation operations such as to count the returned
resultset is also provided via programming languages abstraction.

In order to export the collected data to other data formats, the  command
``mongoexport'' and the developed export script written in Python were used.
The ``mongoexport'', exports the collected data from the database using the
native support to comma-delimited values. As shon in Listing
\ref{file:mongodb-export-command}, the actual execution, the remaining ones
describe how mongo is querying the data from the database. As a result, the
total number of documents exported is shown in on line 8. In addition to the
regular command that exports the entire database, the user can provide a query
string during the execution of the command in order to filter the dataset
returned and, consequently, the number of documents exported. In contrast to
the use of the default export capabilities of mongoDB, an Export script written
in Python, as shown in Listing \ref{file:experiment-export-python}. It uses the
package ``pymongo'' and creates a comma-delimited file with the same header
format and lines as shown in the HTTP Response body of Listing
\ref{file:rtc-ysi-opendap}. The results can be compared with the example of
Listing \ref{file:experiment-export-python-results}.

While the access to the database using the shell command gives full access to 
the CRUD operations, the access to the data was also performed using the REST
Web Services API, as shown in Listing \ref{cmd:mongo-rest-request}. As one can
see, the IP address and the default port number of the mongoDB REST middleware
are provided. Then, the REST operation uses the name of the database and
the collection to correlate with the ones in the database system. In this way,
querying the collections from mongoDB can be performed using the HTTP Request
Parameter variables, by adding the prefix ``filter\underline{ }'' before the
name of the key of the document. The given example shows the query performed in the
collected data for the key ``observation.Conductivity'' with the value
``104.5''. Similarly, the HTTP Request variable ``limit'' is used to reduce
the number of documents returned by the query. Finally, the HTTP Response from
mongoDB returns the result using the JSON format, similar to the ones printed
in the command-shell.

Finally, the access of the collected data was performed using Futon4Mongo,
an open-source web-based application written in Ruby that provides access to the
mongoDB database collections and documents. The screenshots of the session used
to browse through the collection of collected data is shown in Figure
\ref{fig:view-collections-instance-browser-futondb}.

The actual execution of the solution was performed using the demo Mouse
Actions, which consists of a software sensor that perceives any mouse action in
a window area. The experiment consisted of moving the mouse around a white
window, where each action is captured and temporarily saved until a threshold is
reached, making the component to send the messages to the persistence layer, so
that it can save on mongoDB. Figure \ref{fig:experiment-mouseactions-dsp} shows
a screenshot of the moment when the message is sent to mongoDB, being verified
through the HTTP REST Interface using a web browser.

\section{Discussion}

The implementation of the data persistence for NetBEAMS follows the
specifications explained in chapter 4. The purpose of sensor network for
NetBEAMS is to suport data archival from any collected data, and for this
reason, mongoDB was selected after an empirical analysis described in chapter
5. In this way, the requirements described in chapter 4 were expressed as
examples of use cases in section \ref{sec:exp-scenarios}. This chapter
discusses the experiments in different angles: analysis of the taxonomies, the
infrastructure and the use cases evaluation.

\subsection{Taxonomic Evaluation}

Any database system covers the purpose of the sensor data for NetBEAMS, which
is Data Archival, along with the type of storage for the data as an External
Storage. After selecting mongoDB, the development of the persistence layer met
the requirements of archiving data for different types of sensor devices from
the DSP Platform such as the YSI Data Handler and the Mouse Actions demo data.
As it is shown in section \ref{sec:exp-measurements}, the collected data is
archived in files using the BSON format \cite{bson}, a binary representation
used to store instances of the document described in Listing
\ref{file:mongodb-ysi-data-format}. Along with the purpose of Data Archival,
the use of mongoDB also met the specification of an External Storage, since the
data persisted in the file system, as well as Data-Centric for running mongoDB
with different  distributed mongo shards (cluster nodes).

SF-BEAMS uses Tabular Data model, Schema-less data model that represents the
collected data using comma-delimited files, without a database system to manage
those files, just the OPEnDAP middleware. In contrast, the data persistence
proposed by this work proposes mongoDB as a schema-less database system to
manage the collected data collected using NetBEAMS. As described in chapter 5,
the schema-less data model can provide a better support to Data Provenance
because the data model gives support to the change of records without affecting
the others. In this way, the metadata used by the implementation clearly
describes the three fundamental questions regarding the collected data from
sensor devices, as described in section \ref{sec:dsp-persistence-data-model}.
First, the identity of the data is given by the key ``message\underline{
}id'', in a way that it uniquely identifies the data produced by the sensor
identified by the key ``sensor.ip\underline{ }address''. In this way, the
message can be tracked back to the sensor that produced the data for different
purposes such as tracking the data producer. Second, time dimensions were used
in order to track two different aspects of the data collection: the valid time
can be used to correlate observations to the exact moment it occurred, while
the transaction time can be used to verify when the data was collected, and
make decisions such as deleting the instances of data were collected in the
previous two days, etc. Last, but not least, the use of the key
``sensor.location'' can give the exact location of the data, since researchers
can be interested in finding data based on a specific location. Finally, the
most important aspect of the schema-less data model is that the observed data
is added to the key ``observation'', as any sensor device carries its set of
attributes of key-value pairs. A good example of the applicability of the
schema-less model is the scalability regarding changes to the data structure.
For instance, if an existing sensor device is flushed with a new version of
the firmware\footnote{Primary machine instructions that run in the hardware},
changing the format of data types or adding new values, the following instances
of collected data may be different to the previous one by simply adding the
new additions of data. For this reason, the DSP Data Persistence gives
administrators the ability to choose which fields to persist by declaring it
using a bootstrap message for the component, as shown in the parameter
``YSI\underline{ }DESIRED\underline{ }PROPERTIES'' of Listing
\ref{file:dsp-data-persistence-bootstrap.xml}.

In order to access the collected data, the implementation used different APIs
provided by the mongoDB community. The Centralized Query Mechanism was
satisfied by the use of those APIs, which is also fultill the requirements
defined by NetBEAMS. For example, the access of the database using Python or
Java were based on the needs and the use of the languages for different
purposes. In this way, the implementation of the persistence model captures
the instances of any known data collected from NetBEAMS and saves it using the
document pattern as shown in Listing \ref{file:mongodb-ysi-data-format}. Any
programming language used was capable of accessing the same collected data by
their own.

In the end, the implementation of the data persistence for NetBEAMS was
developed in both Single and Distributed Systems. The use of regular single
server execution provided enough computational power for a compatible volume of
data when compared to the one produced from the SF-BEAMS sensor network. In
this way, the implementation suggested by this work fulfill the need of a
persistence system for a Small volume of data being created at the same type
in a Single Server. Furthermore, mongoDB also provided support for Distributed
System, when the use of databse shards, in forms of different mongoDB nodes,
were used as a clustered environment.

\subsection{Infrastructure Evaluation}

Overall, the implementation of the DSP Data Persistence component meets the
specification of the use cases defined in section \ref{sec:exp-scenarios}.
From the perspective of disk-space utilization for the compatible volume of
data for 5 five YSI devices is that it requires a small amount of disk space.
According to the mongoDB documentation \cite{mongodb}, the claimed amount of
disk space is related to how mongoDB uses a binary representation of the JSON
documents, which are composed only of strings, together with the definition of
the indexes. As a consequence, the size of the keys defined for the YSI sonde
data suggests additional disk space were used for the storage.

The performance to insert data into the database system using the Java driver
is documented by mongoDB as the fastest approach, providing persistence
thousands of documents per second. This capacity is more than necessary for
NetBEAMS, which can use the Single System environment. On the other hand,
improvements over the response time can be achieved through the implementation
of a Data-Centric approach on a Distributed Database environment, or clusters.
In this approach, the inserted data is stored in different servers based on a
partition key. Given the fact that less data is concentrated in each
partition, the centralized query mechanism used is faster, and consequently
improves the performance as described in the Data Location taxonomy in chapter
3. In this way, the infrastructural performace during the inserts suggests that
persisting data collected from the NetBEAMS sensors using a clustered
environment is an option when the single server capacity is reached.

Different problems and limitations were faced during the execution of
mongoDB, which were relate to the infrastructure used: the limitations
regarding disk storage in 32bit operating systems and very slow
performance when using virtualized servers. During the execution of the
experiments, mongoDB used memory accordingly, without taking control of the
operating system's resources. However, when mongoDB is used in a 32bit server
architecture, there are limitations on the disk space used to store
collections. In this way, after reaching 2GB of collected data during many
different runs, mongoDB crashed and did not allow the use of the collected
data. For this reason, the setup of a new VirtualBox using Ubuntu Linux 64bits
as a guest to a Mac OS X 64bits was prepared and used throughout the completion
of this work's experiments. On the other hand, the performance was drastically
decreased using mongoDB in virtualized servers due to the fact of dynamic
partition chosen. The performance shell application ``iostat'' was used to
verify the behavior, and the use of mongoDB was switched to the Host operating
system.

\subsection{Use Cases Evaluation}

\cite{sn-programming-language} suggested that programming that programming
languages are easier abstractions to researchers that use collected data from
sensor devices without background in computer science or database systems. The
authors have suggested the implementation of a programming language for this
group of biologists, for instance. However, this suggestion has good
intentions, but limits the users with a single mechanism to access the
collected data from sensors. For this reason, this dissertation proposes the
use of key-value pair databases, such as mongoDB, to provide transparent access
the the data through the abstraction of any programming language that has a
driver written for mongoDB. As one can see during the user experience section,
different programming languages were used to access the collected data, a novel
mechanism not yet reported in the sensor networks community.

The development of the scenarios were clearly written in different ways, such
as the ones defined in Listing \ref{file:experiment-scenarios}, which
implements the use cases for data retrieval, update and deletion. The main
advantage of this approach is that users do not use the SQL language to extract
data from the database. Rather, the use of the method calls defined for each
operation, as well as the concept of dynamic binding and objects, gives simple,
yet powerful syntax to access data. For example, based on the description
of the collected data by keys and values shown in Listing
\ref{file:mongodb-ysi-data-format}, different use cases can evaluate documents
in this structure. Usually researchers may search the collected data by a given
range of dates, for their regular activities of data analysis. The scenario
implemented by R1 in the abovecited listing shows it clearly. Similarly, the
search for data in a given sensor device is expressed by the search for the IP
address for them on the use cased defined by R2. Finally, the search for the
raw data obtained by the sensor devices is used from the key ``observation'',
as shown by the R3 use case implementation.

Accompanying that non-predictable events might occur during the data
collection process of sensor devices, and consequently, influence in their
values, the support to annotations is a feature that can be used to describe
the data. mongoDB indirectly supports the use of annotation because of its free
method of updating documents and adding new keys to any of them without
requiring design changes. Supposing an environmental accident happens, such as
the San Francisco oil spill of October 2009 \cite{sfbay-oilspill2009}, the
values of the collected data, and consequently data analysis, be influenced by
different ranges of values. In this way, the implementation of the scenario U1
provided the implementation of the addition of a new key called ``tag'', with
the annotation ``oil spill''. This new key, after properly indexed, provides
another option to meaningfully search the collected data from the sensor
devices. Last, but not least, old data can be deleted as expressed by D1, as an
operation supported by any database system.

One common approach during research is data sharing. Researchers can choose
the use of spreadsheets to analyze collected data with simpler tools than
database systems or programming language. For this reason, OPEnDAP is used in
the community to share data using a comma-delited format. In this way, the
implementation of the use cases for R4 were developed in two different ways.
First, the use of the export capabilities to a common-delimited values was
performed \ref{file:mongodb-export-command}, taking approximaly 3.3 minutes to
export one year worth of data for 5 YSI devices without indexing. Second, the
implementation of a custom export utility using Python, as shown in Listing
\ref{file:experiment-export-python} enabled the creation of the comma-delimited
files with the same format as the ones used by OPEnDAP. The comparison can be
done by analyzing the output of the OPEnDAP request from SF-BEAMS on Listing
\ref{file:rtc-ysi-opendap}, and the output provided from the script written in
Python, show in Listing \ref{file:experiment-export-python-results}.
Furthermore, considering research teams with skills in Web Services, an easy
ubiquos integration with mongoDB can be provided by using the REST Web Services
APIs provided by mongoDB, still in alpha version. A practical application of
this API was the created Apache Futon 4 mongoDB web application, used to give
user friendlier access to the collected data as shown in 
\ref{fig:view-collections-instance-browser-futondb}.

As discussed in chapter 5, the differences between a traditional database
system, which uses the relational data model, and a document-oriented database
system were clearly noticed during the development of this work. As mentioned
in the section \ref{sec:experiments-user-experimence}, the user experience was
mostly related to the definition of the keys that describe the sensor device,
as discussed during the design of the solution in chapter 6. On the contraty,
the abstraction of objects and collections during the data access using
different programming languages. Therefore, based on the use of a different
type of data model in the scope of data persistence for sensor networks, the
taxonomies defined in chapter 3 could be updated to the one shown in Figure
\ref{fig:taxonomy-data-model-modified}, including the Key-Value Pair Data
Model.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-model-modified}
  \caption{Data Model Taxonomy}
  \label{fig:taxonomy-data-model-modified}
\end{figure}

\subsection{Difficulties Experienced}
\label{sec:experiments-difficulties}

The inception of the experiments described in by this work moments of Software
Engineering. Particularly speaking about the learning curve, different
situations contributed to the success of the first experiments. Since the
technoogy is new, it brought new challenges regarding the usability and user
experience. The technology is relatively new and still under development,
which many different bugs were be reproduced by random during the execution of
the experiments, and therefore, reports were sent to the open-source community
through use different available communication channels. First, the fastest way
to solve an issue was through the mongoDB's IRC channel located at the Freenode
server (www.freenode.org, channel mongodb). Second, the use of the users
group, located at http://groups.google.com/group/mongodb-user also helped
solving problems.

Physical memory is limited and, for this reason, it may not be enough to run
the experiments for total workloads. In this case, the Java Virtual Memory was
totally consumed by one of the revisions of the experiments setup. Therefore, 
it was learned that large amount of data loads require different approach to
manage data in memory accordingly, such as creating different object pools as
temporary cache for objects whose state are similar to others.

Also related to the difficulties of this work, the fact that mongoDB is not yet
stable makes the implementation a bit harder. The main mongod process is on its
1.1.4 version as the time this line is writte, while the shards support is
still in alpha version. For instance, different work-arounds were performed to
have shards experimented as described, but the execution was not successful.

To sum up, this chapter described the experimental results of the
implementation of a data persistence component for NetBEAMS, showing the
experiment results, the measurement results and finally the discussion about
the solution in general. The following section brings the conclusions and
future works for this dissertation.