### 2009-12-08 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Conclusions and Future Works}

There were important important remarks covered in this work, including the
importance that sensor networks are becoming ubiquitous, having their produced
data available online and other types of media. In this way, in order to
provide access to the collected data, one must first assess the different
types of infrastructural characteristics of the studied sensor networks. In
general, the selection of a technology that provides such a persistence
solution capabilities can  part of a process of analysis of taxonomies
proposed by this work, characterized as the first contribution of this
disseration.

In my own opinion, based on the findings of the taxonomies, together with
the empirical analysis of the technologies, this work revealed a novel approach
to provide persistence and access to collected data from sensor devices using
the Key-Value-Pair data model. Besides the contribution of a new component
component for the Data Sensor Platform, the evidence that the system improves
productivity was confirmed during the execution of the experiments, writing
less source-code to obtain better functionalities when compared to other
solutions. Similarly, improved work suggests that the abstraction of
programming language should be used with research groups whose primary
expertise is not in Computer Science or Database Systems.

All in all, the DSP Data Persistence component implemented for the NetBEAMS
solution to SF-BEAMS can be used in any network server. The solution for data
persistence is a novice way to save data into the database, provided by
mongoDB's ability to cope with the uncertainty of any properties collected from
any sensor specification.

As far as future works, there are many different directions for future works.
First, the first major release of mongoDB with complete support to shards is
scheduled for 2010. An important mechanism of mongoDB is the fact that it can
manage different mongoDB servers spread as a cluster, as it gives support to
the Data-Centric Storage approach. In this way, this approach can reduce the
retrieval time by focusing on centralized selection of data over a given
database shard. Moreover, the overall search through the collected data
attributes can be drastically decreased by parallel search over the database
shards using the MapReduce \cite{map-reduce} tools. At the time of development
of the experiments, the MapReduce API from mongoDB was also in the alpha
version, and therefore, presenting lots of bugs. one suggestion for the
implementation of such functionality is based in Figure
\ref{fig:future-works-data-centric-map-reduce}, which shows an architectural
idea of the deployment of a Data-Centric storage approach. When searching for
specific document attributes, the process is sent in parallel to each of the
database shards, and the result is collected by the cluster head. After
calculating the final result, mongoDB server returns the relevant result to
the user. I was personally expecting to implement the scario of data retrieval
of collected daa.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/future-works-data-centric-map-reduce}
  \caption{Data-Centric approach using MapReduce to search accross different
  database shards}
  \label{fig:future-works-data-centric-map-reduce}
\end{figure}

In addition to infrastructural, the use techniques for scheduling data
gathering from the SF-BEAMS sensor networks can be used to decrease the data
load into the system. While the period of times used by the RTC seems
to cover the basic requirements of data, the data storage may be
seen as overloaded with repeated data. Since environmental conditions
does not change drastically over time, the use of data scheduling and
data clustering approaches can be used. First, considering that SF-BEAMS
contains a cluster of sensors, the data load could be decreased by using
techniques such as clustering the data in the data sink before they are
persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
the knowledge about the data can only be achieved in the data sink. In this
way, one suggestion is the implementation of a DSP Data Clustering component
that is responsible for the data clustering before they are inserted into the
database. Similarly, the use of schedulers on the sensors could also be used to
decrease the amount of data to be sent to the data sink, as seen at the survey
\cite{sn-scheduling-survey}.

Finally, taking into account the use of the database system to complement
NetBEAMS execution, the design of event-based applications could add different
functionalities to the management capabilities of the sensor networks. For
instance, the YSI sonde collected data carries the information regarding the
battery life of the device, seen at the collected attribute 
``observation.Battery''. A new DSP Component responsible for observing these
attributes could define the threshold of the event of low-battery. With an
event-based application in mind, events regarding the SF-BEAMS network could be
better managed through NetBEAMS where scheduled sites visits would depend on
such events to happen and, thus, avoiding unecessary operational costs.