### 2009-12-12 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Conclusions and Future Works}

There were significant remarks covered in this work, including the
importance that sensor networks are becoming ubiquitous, having their produced
data available online and other types of media. In order to provide access to
the collected data, one must first assess the different types of
infrastructural characteristics of the studied sensor networks. The selection
of a technology that provides such a persistence solution capabilities can be 
part of a process of analysis of taxonomies proposed by this work,
characterized as the first contribution of this report.

Based on the findings of the taxonomies, together with the empirical analysis
of the technologies, this work revealed a novel approach to provide
persistence and access to collected data from sensor devices using the
Key-Value-Pair data model. Besides the contribution of a new component for the
Data Sensor Platform, the evidence that the system improves productivity was
confirmed during the execution of the experiments, writing less source-code to
obtain better functionalities when compared to other solutions. Similarly,
improved work suggests that the abstraction of programming language should be
used with research groups whose primary expertise is not in Computer Science
or Database Systems.

The DSP Data Persistence component improves NetBEAMS's support to SF-BEAMS,
incorporating a persistence layer for the sensor network that can be used in
different ways. The solution for data persistence using Key-Value Pair
databases is a novel way to save collected data from sensor devices into a
single or multiple servers of a database technology. mongoDB's ability to cope
with the uncertainty of any properties collected from any sensor specification
is the most compeling reason to use the Key-value pair data model for collected
data from sensor devices.

As far as future works, there are many different directions for exploration.
The first major release of mongoDB with complete support to shards is
scheduled for 2010. An important mechanism of mongoDB is that it can
manage different mongoDB servers spread out in a cluster. This feature will give
better support to the Data-Centric Storage approach. This approach can reduce
the retrieval time by focusing on centralized selection of data over a given
database shard. Moreover, the overall search through the collected data
attributes can be drastically decreased by a parallel search over the database
shards using the MapReduce \cite{map-reduce} tools. At the time of development
of the experiments, the MapReduce API from mongoDB was also in the alpha
version, and therefore, presented many bugs. One suggestion for the
implementation of such functionality is based in Figure
\ref{fig:future-works-data-centric-map-reduce}, which shows an architectural
idea of the deployment of a Data-Centric storage approach. When searching for
specific document attributes, the process is sent in parallel to each of the
database shards, and the cluster head collects the result. After calculating
the final result, mongoDB server returns the relevant result to the user.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.65]{../diagrams/future-works-data-centric-map-reduce}
  \caption{Data-Centric approach using MapReduce to search across different
  database shards}
  \label{fig:future-works-data-centric-map-reduce}
\end{figure}

In addition to the infrastructure, the use techniques for scheduling data
gathering from the SF-BEAMS sensor networks can be used to decrease the data
load into the system. While the period of times used by RTC seems
to cover the basic requirements of data, the data storage may be
seen as overloaded with repeated data. Since environmental conditions
might not change drastically over time, the use of data scheduling and
data clustering approaches can be used. Considering that SF-BEAMS
contains a cluster of sensors, the data load could be decreased by using
techniques such as clustering the data in the data sink before they are
persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
the knowledge about the data can only be achieved in the data sink. One
suggestion is the implementation of a DSP Data Clustering component that is
responsible for the data clustering before they are inserted into the
database. Similarly, the use of schedulers on the sensors could also be used
to decrease the amount of data to be sent to the data sink, as seen at the
survey \cite{sn-scheduling-survey}.

Finally, taking into account the use of the database system to complement
NetBEAMS execution, the design of event-based applications could add different
functionalities to the management capabilities of the sensor networks. For
instance, the collected data from a YSI device \cite{YSI-Sonde} carries the
information regarding the battery life of the device, seen at the collected
attribute ``observation.Battery''. A new DSP Component responsible for
observing these attributes could define the threshold of the event of
low-battery. With an event-based application in mind, events regarding the
SF-BEAMS network could be better managed through NetBEAMS where scheduled
sites visits would depend on such events to happen and, thus, avoiding
unnecessary operational costs.